### 2020-04-28 figure2.tex

% Created by tikzDevice version 0.12.3 on 2020-04-27 21:17:30
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 37.26, 33.48) rectangle (283.08,264.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 37.26, 33.48) rectangle (283.08,264.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
	(283.08, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 77.88) --
	(283.08, 77.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,122.28) --
	(283.08,122.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,166.68) --
	(283.08,166.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,211.08) --
	(283.08,211.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,255.48) --
	(283.08,255.48);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 65.62,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (112.90,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (160.17,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (207.44,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.72,119.58) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
	( 65.62,122.88);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
	( 65.62,122.88);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
	( 65.62,122.88);

\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90,234.50) --
	(112.90,234.50);

\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90,234.50) --
	(112.90, 55.99);

\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90, 55.99) --
	(112.90, 55.99);

\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17,153.48) --
	(160.17,153.48);

\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17,153.48) --
	(160.17, 99.27);

\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17, 99.27) --
	(160.17, 99.27);

\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44,164.15) --
	(207.44,164.15);

\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44,164.15) --
	(207.44, 68.61);

\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44, 68.61) --
	(207.44, 68.61);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72,260.46) --
	(254.72,260.46);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72,260.46) --
	(254.72, 41.21);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72, 41.21) --
	(254.72, 41.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
	( 37.26,264.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86, 30.17) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86, 74.57) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,118.97) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,163.37) {3.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,207.78) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,252.18) {4.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26, 33.48) --
	( 37.26, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26, 77.88) --
	( 37.26, 77.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,122.28) --
	( 37.26,122.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,166.68) --
	( 37.26,166.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,211.08) --
	( 37.26,211.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,255.48) --
	( 37.26,255.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
	(283.08, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 65.62, 30.48) --
	( 65.62, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.90, 30.48) --
	(112.90, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (160.17, 30.48) --
	(160.17, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.44, 30.48) --
	(207.44, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.72, 30.48) --
	(254.72, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 65.62, 21.46) {base};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (112.90, 21.46) {$\mu_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (160.17, 21.46) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (207.44, 21.46) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (254.72, 21.46) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (160.17,  8.33) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 14.26,148.92) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.44] at ( 37.26,273.16) {A. Baseline};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] (326.34, 33.48) rectangle (572.16,264.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (326.34, 33.48) rectangle (572.16,264.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 79.65) --
	(572.16, 79.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34,141.22) --
	(572.16,141.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34,202.79) --
	(572.16,202.79);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (354.70,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (401.98,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (449.25,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (496.52,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (543.80,140.42) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
	(354.70,142.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
	(354.70,142.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
	(354.70,142.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,187.46) --
	(401.98,187.46);

\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,187.46) --
	(401.98,111.70);

\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,111.70) --
	(401.98,111.70);

\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,195.31) --
	(449.25,195.31);

\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,195.31) --
	(449.25,101.35);

\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,101.35) --
	(449.25,101.35);

\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52,213.80) --
	(496.52,213.80);

\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52,213.80) --
	(496.52, 48.20);

\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52, 48.20) --
	(496.52, 48.20);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80,261.16) --
	(543.80,261.16);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80,261.16) --
	(543.80, 41.24);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80, 41.24) --
	(543.80, 41.24);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 33.48) --
	(326.34,264.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94, 76.35) {2.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94,137.92) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94,199.49) {3.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34, 79.65) --
	(326.34, 79.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34,141.22) --
	(326.34,141.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34,202.79) --
	(326.34,202.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 33.48) --
	(572.16, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.70, 30.48) --
	(354.70, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.98, 30.48) --
	(401.98, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (449.25, 30.48) --
	(449.25, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.52, 30.48) --
	(496.52, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (543.80, 30.48) --
	(543.80, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (354.70, 21.46) {base};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (401.98, 21.46) {$\hat{\mu}_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (449.25, 21.46) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (496.52, 21.46) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (543.80, 21.46) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (449.25,  8.33) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (303.34,148.92) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.44] at (326.34,273.16) {B. Reduced uncertainty in $r$};
\end{scope}
\end{tikzpicture}

\end{document}