### 2020-05-07 compare_assumption.tex
done

% Created by tikzDevice version 0.12.3 on 2020-05-07 17:10:16
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{gray}{0.95}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (  0.00,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.04,  0.00) -- ( 38.04,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07,  0.00) -- ( 76.07,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.11,  0.00) -- (114.11,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.15,  0.00) -- (152.15,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.18,  0.00) -- (190.18,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.22,  0.00) -- (228.22,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.26,  0.00) -- (266.26,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.29,  0.00) -- (304.29,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.33,  0.00) -- (342.33,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (380.37,  0.00) -- (380.37,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.41,  0.00) -- (418.41,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.44,  0.00) -- (456.44,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.48,  0.00) -- (494.48,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (532.52,  0.00) -- (532.52,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.55,  0.00) -- (570.55,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (608.59,  0.00) -- (608.59,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (646.63,  0.00) -- (646.63,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (684.66,  0.00) -- (684.66,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (722.70,  0.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 11.41) -- (722.70, 11.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 22.82) -- (722.70, 22.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 34.23) -- (722.70, 34.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 45.64) -- (722.70, 45.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 57.06) -- (722.70, 57.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 68.47) -- (722.70, 68.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 79.88) -- (722.70, 79.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 91.29) -- (722.70, 91.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,102.70) -- (722.70,102.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,114.11) -- (722.70,114.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,125.52) -- (722.70,125.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,136.93) -- (722.70,136.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,148.34) -- (722.70,148.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,159.75) -- (722.70,159.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,171.17) -- (722.70,171.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,182.58) -- (722.70,182.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,193.99) -- (722.70,193.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,205.40) -- (722.70,205.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,216.81) -- (722.70,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (240.90,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (240.90,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 71.72, 37.19) rectangle (235.40,211.31);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 71.72, 37.19) rectangle (235.40,211.31);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 93.56, 37.19) --
	( 93.56,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (131.56, 37.19) --
	(131.56,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (169.55, 37.19) --
	(169.55,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (207.55, 37.19) --
	(207.55,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72, 49.93) --
	(235.40, 49.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72, 71.16) --
	(235.40, 71.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72, 92.40) --
	(235.40, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,113.63) --
	(235.40,113.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,134.87) --
	(235.40,134.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,156.10) --
	(235.40,156.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,177.34) --
	(235.40,177.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,198.57) --
	(235.40,198.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 74.56, 37.19) --
	( 74.56,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.56, 37.19) --
	(112.56,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.55, 37.19) --
	(150.55,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.55, 37.19) --
	(188.55,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.55, 37.19) --
	(226.55,211.31);
\definecolor{drawColor}{RGB}{255,165,0}
\definecolor{fillColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (131.47, 71.16) circle (  2.50);
\definecolor{drawColor}{RGB}{160,32,240}
\definecolor{fillColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (128.69, 92.40) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (128.47,113.63) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (117.88,198.57) circle (  2.50);
\definecolor{drawColor}{RGB}{160,32,240}
\definecolor{fillColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (169.46,134.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (108.06,156.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.22,177.34) circle (  2.50);
\definecolor{drawColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (167.58, 71.16) --
	(167.58, 71.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] (167.58, 71.16) --
	( 95.51, 71.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 95.51, 71.16) --
	( 95.51, 71.16);
\definecolor{drawColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.8pt,line join=round] (163.42, 92.40) --
	(163.42, 92.40);

\path[draw=drawColor,line width= 0.8pt,line join=round] (163.42, 92.40) --
	( 96.55, 92.40);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 96.55, 92.40) --
	( 96.55, 92.40);

\path[draw=drawColor,line width= 0.8pt,line join=round] (141.83,113.63) --
	(141.83,113.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (141.83,113.63) --
	(114.86,113.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (114.86,113.63) --
	(114.86,113.63);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) --
	(117.88,198.57);

\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) --
	(117.88,198.57);

\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) --
	(117.88,198.57);
\definecolor{drawColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.8pt,line join=round] (174.59,134.87) --
	(174.59,134.87);

\path[draw=drawColor,line width= 0.8pt,line join=round] (174.59,134.87) --
	(164.38,134.87);

\path[draw=drawColor,line width= 0.8pt,line join=round] (164.38,134.87) --
	(164.38,134.87);

\path[draw=drawColor,line width= 0.8pt,line join=round] (149.69,156.10) --
	(149.69,156.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] (149.69,156.10) --
	( 79.16,156.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 79.16,156.10) --
	( 79.16,156.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] (187.73,177.34) --
	(187.73,177.34);

\path[draw=drawColor,line width= 0.8pt,line join=round] (187.73,177.34) --
	(152.24,177.34);

\path[draw=drawColor,line width= 0.8pt,line join=round] (152.24,177.34) --
	(152.24,177.34);
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (105.33, 49.93) --
	(105.57, 49.93) --
	(105.81, 49.93) --
	(106.05, 49.93) --
	(106.29, 49.93) --
	(106.53, 49.93) --
	(106.77, 49.93) --
	(107.01, 49.93) --
	(107.25, 49.93) --
	(107.49, 49.93) --
	(107.73, 49.93) --
	(107.97, 49.93) --
	(108.21, 49.93) --
	(108.45, 49.93) --
	(108.69, 49.93) --
	(108.93, 49.93) --
	(109.17, 49.93) --
	(109.41, 49.93) --
	(109.65, 49.93) --
	(109.89, 49.93) --
	(110.13, 49.93) --
	(110.37, 49.93) --
	(110.61, 49.93) --
	(110.85, 49.93) --
	(111.09, 49.93) --
	(111.33, 49.93) --
	(111.57, 49.93) --
	(111.81, 49.93) --
	(112.05, 49.93) --
	(112.29, 49.93) --
	(112.53, 49.93) --
	(112.77, 49.93) --
	(113.01, 49.93) --
	(113.25, 49.93) --
	(113.49, 49.93) --
	(113.73, 49.93) --
	(113.97, 49.93) --
	(114.21, 49.93) --
	(114.45, 49.93) --
	(114.69, 49.93) --
	(114.93, 49.93) --
	(115.17, 49.93) --
	(115.41, 49.93) --
	(115.65, 49.93) --
	(115.89, 49.93) --
	(116.13, 49.93) --
	(116.37, 49.93) --
	(116.61, 49.93) --
	(116.85, 49.93) --
	(117.09, 49.93) --
	(117.33, 49.93) --
	(117.57, 49.93) --
	(117.81, 49.93) --
	(118.05, 49.93) --
	(118.29, 49.93) --
	(118.53, 49.93) --
	(118.77, 49.93) --
	(119.01, 49.93) --
	(119.25, 49.93) --
	(119.49, 49.93) --
	(119.73, 49.93) --
	(119.97, 49.93) --
	(120.21, 49.93) --
	(120.44, 49.93) --
	(120.68, 49.93) --
	(120.92, 49.93) --
	(121.16, 49.93) --
	(121.40, 49.93) --
	(121.64, 49.93) --
	(121.88, 49.93) --
	(122.12, 49.93) --
	(122.36, 49.93) --
	(122.60, 49.93) --
	(122.84, 49.93) --
	(123.08, 49.93) --
	(123.32, 49.93) --
	(123.56, 49.93) --
	(123.80, 49.93) --
	(124.04, 49.93) --
	(124.28, 49.93) --
	(124.52, 49.93) --
	(124.76, 49.93) --
	(125.00, 49.93) --
	(125.24, 49.93) --
	(125.48, 49.93) --
	(125.72, 49.93) --
	(125.96, 49.93) --
	(126.20, 49.93) --
	(126.44, 49.93) --
	(126.68, 49.93) --
	(126.92, 49.93) --
	(127.16, 49.93) --
	(127.40, 49.93) --
	(127.64, 49.93) --
	(127.88, 49.93) --
	(128.12, 49.93) --
	(128.36, 49.93) --
	(128.60, 49.93) --
	(128.84, 49.93) --
	(129.08, 49.93) --
	(129.32, 49.93) --
	(129.56, 49.93) --
	(129.80, 49.93) --
	(130.04, 49.93) --
	(130.28, 49.93) --
	(130.52, 49.93) --
	(130.76, 49.93) --
	(131.00, 49.93) --
	(131.24, 49.93) --
	(131.48, 49.93) --
	(131.72, 49.93) --
	(131.96, 49.93) --
	(132.20, 49.93) --
	(132.44, 49.93) --
	(132.68, 49.93) --
	(132.92, 49.93) --
	(133.16, 49.93) --
	(133.40, 49.93) --
	(133.64, 49.93) --
	(133.88, 49.93) --
	(134.12, 49.93) --
	(134.36, 49.93) --
	(134.60, 49.93) --
	(134.84, 49.93) --
	(135.08, 49.93) --
	(135.32, 49.93) --
	(135.56, 49.93) --
	(135.80, 49.93) --
	(136.04, 49.93) --
	(136.28, 49.93) --
	(136.52, 49.93) --
	(136.76, 49.93) --
	(137.00, 49.93) --
	(137.24, 49.93) --
	(137.48, 49.93) --
	(137.72, 49.93) --
	(137.96, 49.93) --
	(138.20, 49.93) --
	(138.44, 49.93) --
	(138.68, 49.93) --
	(138.92, 49.93) --
	(139.16, 49.93) --
	(139.40, 49.93) --
	(139.64, 49.93) --
	(139.88, 49.93) --
	(140.12, 49.93) --
	(140.36, 49.93) --
	(140.60, 49.93) --
	(140.84, 49.93) --
	(141.08, 49.93) --
	(141.32, 49.93) --
	(141.56, 49.93) --
	(141.80, 49.93) --
	(142.04, 49.93) --
	(142.28, 49.93) --
	(142.52, 49.93) --
	(142.76, 49.93) --
	(143.00, 49.93) --
	(143.24, 49.93) --
	(143.48, 49.93) --
	(143.72, 49.93) --
	(143.96, 49.93) --
	(144.20, 49.93) --
	(144.44, 49.93) --
	(144.68, 49.93) --
	(144.92, 49.93) --
	(145.16, 49.93) --
	(145.40, 49.93) --
	(145.64, 49.93) --
	(145.88, 49.93) --
	(146.12, 49.93) --
	(146.36, 49.93) --
	(146.60, 49.93) --
	(146.84, 49.93) --
	(147.08, 49.93) --
	(147.32, 49.93) --
	(147.56, 49.93) --
	(147.80, 49.93) --
	(148.04, 49.93) --
	(148.28, 49.93) --
	(148.52, 49.93) --
	(148.76, 49.93) --
	(149.00, 49.93) --
	(149.24, 49.93) --
	(149.48, 49.93) --
	(149.72, 49.93) --
	(149.96, 49.93) --
	(150.20, 49.93) --
	(150.44, 49.93) --
	(150.68, 49.93) --
	(150.92, 49.93) --
	(151.16, 49.93) --
	(151.40, 49.93) --
	(151.64, 49.93) --
	(151.88, 49.93) --
	(152.12, 49.93) --
	(152.36, 49.93) --
	(152.60, 49.93) --
	(152.84, 49.93) --
	(153.08, 49.93) --
	(153.32, 49.93) --
	(153.56, 49.93) --
	(153.80, 49.93) --
	(154.04, 49.93) --
	(154.28, 49.93) --
	(154.52, 49.93) --
	(154.76, 49.93) --
	(155.00, 49.93) --
	(155.24, 49.93) --
	(155.48, 49.93) --
	(155.72, 49.93) --
	(155.96, 49.93) --
	(156.20, 49.93) --
	(156.44, 49.93) --
	(156.68, 49.93) --
	(156.92, 49.93) --
	(157.16, 49.93) --
	(157.40, 49.93) --
	(157.64, 49.93) --
	(157.88, 49.93) --
	(158.12, 49.93) --
	(158.36, 49.93) --
	(158.60, 49.93) --
	(158.84, 49.93) --
	(159.08, 49.93) --
	(159.32, 49.93) --
	(159.56, 49.93) --
	(159.80, 49.93) --
	(160.04, 49.93) --
	(160.28, 49.93) --
	(160.52, 49.93) --
	(160.76, 49.93) --
	(161.00, 49.93) --
	(161.24, 49.93) --
	(161.48, 49.93) --
	(161.72, 49.93) --
	(161.96, 49.93) --
	(162.20, 49.93) --
	(162.44, 49.93) --
	(162.68, 49.93) --
	(162.92, 49.93) --
	(163.16, 49.93) --
	(163.40, 49.93) --
	(163.64, 49.93) --
	(163.88, 49.93) --
	(164.12, 49.93) --
	(164.36, 49.93) --
	(164.60, 49.93) --
	(164.84, 49.93) --
	(165.08, 49.93) --
	(165.32, 49.93) --
	(165.56, 49.93) --
	(165.80, 49.93) --
	(166.04, 49.93) --
	(166.28, 49.93) --
	(166.52, 49.93) --
	(166.76, 49.93) --
	(167.00, 49.93) --
	(167.24, 49.93) --
	(167.48, 49.93) --
	(167.72, 49.93) --
	(167.96, 49.93) --
	(168.20, 49.93) --
	(168.44, 49.93) --
	(168.68, 49.93) --
	(168.92, 49.93) --
	(169.16, 49.93) --
	(169.40, 49.93) --
	(169.64, 49.93) --
	(169.88, 49.93) --
	(170.12, 49.93) --
	(170.36, 49.93) --
	(170.60, 49.93) --
	(170.84, 49.93) --
	(171.08, 49.93) --
	(171.32, 49.93) --
	(171.56, 49.93) --
	(171.80, 49.93) --
	(172.04, 49.93) --
	(172.28, 49.93) --
	(172.52, 49.93) --
	(172.76, 49.93) --
	(173.00, 49.93) --
	(173.24, 49.93) --
	(173.48, 49.93) --
	(173.72, 49.93) --
	(173.96, 49.93) --
	(174.20, 49.93) --
	(174.44, 49.93) --
	(174.68, 49.93) --
	(174.92, 49.93) --
	(175.16, 49.93) --
	(175.40, 49.93) --
	(175.64, 49.93) --
	(175.88, 49.93) --
	(176.12, 49.93) --
	(176.36, 49.93) --
	(176.60, 49.93) --
	(176.84, 49.93) --
	(177.08, 49.93) --
	(177.32, 49.93) --
	(177.56, 49.93) --
	(177.80, 49.93) --
	(178.04, 49.93) --
	(178.28, 49.93) --
	(178.52, 49.93) --
	(178.76, 49.93) --
	(179.00, 49.93) --
	(179.24, 49.93) --
	(179.48, 49.93) --
	(179.72, 49.93) --
	(179.96, 49.93) --
	(180.20, 49.93) --
	(180.44, 49.93) --
	(180.68, 49.93) --
	(180.92, 49.93) --
	(181.16, 49.93) --
	(181.40, 49.93) --
	(181.64, 49.93) --
	(181.88, 49.93) --
	(182.12, 49.93) --
	(182.36, 49.93) --
	(182.60, 49.93) --
	(182.84, 49.93) --
	(183.08, 49.93) --
	(183.32, 49.93) --
	(183.56, 49.93) --
	(183.80, 49.93) --
	(184.04, 49.93) --
	(184.28, 49.93) --
	(184.52, 49.93) --
	(184.76, 49.93) --
	(185.00, 49.93) --
	(185.24, 49.93) --
	(185.48, 49.93) --
	(185.72, 49.93) --
	(185.96, 49.93) --
	(186.20, 49.93) --
	(186.44, 49.93) --
	(186.68, 49.93) --
	(186.92, 49.93) --
	(187.16, 49.93) --
	(187.40, 49.93) --
	(187.64, 49.93) --
	(187.88, 49.93) --
	(188.12, 49.93) --
	(188.36, 49.93) --
	(188.60, 49.93) --
	(188.84, 49.93) --
	(189.08, 49.93) --
	(189.32, 49.93) --
	(189.56, 49.93) --
	(189.80, 49.93) --
	(190.04, 49.93) --
	(190.28, 49.93) --
	(190.52, 49.93) --
	(190.76, 49.93) --
	(191.00, 49.93) --
	(191.24, 49.93) --
	(191.48, 49.93) --
	(191.72, 49.93) --
	(191.96, 49.93) --
	(192.20, 49.93) --
	(192.44, 49.93) --
	(192.68, 49.93) --
	(192.92, 49.93) --
	(193.16, 49.93) --
	(193.40, 49.93) --
	(193.64, 49.93) --
	(193.88, 49.93) --
	(194.12, 49.93) --
	(194.36, 49.93) --
	(194.60, 49.93) --
	(194.84, 49.93) --
	(195.08, 49.93) --
	(195.32, 49.93) --
	(195.56, 49.93) --
	(195.80, 49.93) --
	(196.04, 49.93) --
	(196.28, 49.93) --
	(196.52, 49.93) --
	(196.76, 49.93) --
	(197.00, 49.93) --
	(197.24, 49.93) --
	(197.48, 49.93) --
	(197.72, 49.93) --
	(197.96, 49.93) --
	(198.20, 49.93) --
	(198.44, 49.93) --
	(198.68, 49.93) --
	(198.92, 49.93) --
	(199.16, 49.93) --
	(199.40, 49.93) --
	(199.64, 49.93) --
	(199.88, 49.93) --
	(200.12, 49.93) --
	(200.36, 49.93) --
	(200.60, 49.93) --
	(200.84, 49.93) --
	(201.08, 49.93) --
	(201.32, 49.93) --
	(201.56, 49.93) --
	(201.80, 49.93) --
	(202.04, 49.93) --
	(202.28, 49.93) --
	(202.52, 49.93) --
	(202.76, 49.93) --
	(203.00, 49.93) --
	(203.24, 49.93) --
	(203.48, 49.93) --
	(203.72, 49.93) --
	(203.96, 49.93) --
	(204.20, 49.93) --
	(204.44, 49.93) --
	(204.68, 49.93) --
	(204.92, 49.93) --
	(205.16, 49.93) --
	(205.40, 49.93) --
	(205.64, 49.93) --
	(205.88, 49.93) --
	(206.12, 49.93) --
	(206.36, 49.93) --
	(206.60, 49.93) --
	(206.84, 49.93) --
	(207.08, 49.93) --
	(207.32, 49.93) --
	(207.56, 49.93) --
	(207.80, 49.93) --
	(208.04, 49.93) --
	(208.28, 49.93) --
	(208.52, 49.93) --
	(208.76, 49.93) --
	(209.00, 49.93) --
	(209.24, 49.93) --
	(209.48, 49.93) --
	(209.72, 49.93) --
	(209.96, 49.93) --
	(210.20, 49.93) --
	(210.44, 49.93) --
	(210.68, 49.93) --
	(210.92, 49.93) --
	(211.16, 49.93) --
	(211.40, 49.93) --
	(211.64, 49.93) --
	(211.88, 49.93) --
	(212.12, 49.93) --
	(212.36, 49.93) --
	(212.60, 49.93) --
	(212.84, 49.93) --
	(213.08, 49.93) --
	(213.32, 49.93) --
	(213.56, 49.93) --
	(213.80, 49.93) --
	(214.04, 49.93) --
	(214.28, 49.93) --
	(214.52, 49.93) --
	(214.76, 49.93) --
	(215.00, 49.93) --
	(215.24, 49.93) --
	(215.48, 49.93) --
	(215.72, 49.93) --
	(215.96, 49.93) --
	(216.20, 49.93) --
	(216.44, 49.93) --
	(216.68, 49.93) --
	(216.92, 49.93) --
	(217.16, 49.93) --
	(217.40, 49.93) --
	(217.64, 49.93) --
	(217.88, 49.93) --
	(218.12, 49.93) --
	(218.36, 49.93) --
	(218.60, 49.93) --
	(218.84, 49.93) --
	(219.08, 49.93) --
	(219.32, 49.93) --
	(219.56, 49.93) --
	(219.80, 49.93) --
	(220.04, 49.93) --
	(220.28, 49.93) --
	(220.52, 49.93) --
	(220.76, 49.93) --
	(221.00, 49.93) --
	(221.24, 49.93) --
	(221.48, 49.93) --
	(221.72, 49.93) --
	(221.96, 49.93) --
	(222.20, 49.93) --
	(222.44, 49.93) --
	(222.68, 49.93) --
	(222.92, 49.93) --
	(223.16, 49.93) --
	(223.40, 49.93) --
	(223.64, 49.93) --
	(223.88, 49.93) --
	(224.12, 49.93) --
	(224.36, 49.93) --
	(224.60, 49.93) --
	(224.84, 49.93) --
	(225.08, 49.93) --
	(225.32, 49.93) --
	(225.56, 49.93) --
	(225.80, 49.93) --
	(226.04, 49.93) --
	(226.28, 49.93) --
	(226.52, 49.93) --
	(226.76, 49.93) --
	(227.00, 49.93) --
	(227.24, 49.93) --
	(227.48, 49.93) --
	(227.72, 49.93) --
	(227.96, 49.93) --
	(227.96, 49.96) --
	(227.72, 49.96) --
	(227.48, 49.96) --
	(227.24, 49.95) --
	(227.00, 49.95) --
	(226.76, 49.95) --
	(226.52, 49.95) --
	(226.28, 49.95) --
	(226.04, 49.95) --
	(225.80, 49.95) --
	(225.56, 49.95) --
	(225.32, 49.94) --
	(225.08, 49.94) --
	(224.84, 49.94) --
	(224.60, 49.94) --
	(224.36, 49.94) --
	(224.12, 49.94) --
	(223.88, 49.94) --
	(223.64, 49.93) --
	(223.40, 49.93) --
	(223.16, 49.93) --
	(222.92, 49.93) --
	(222.68, 49.93) --
	(222.44, 49.93) --
	(222.20, 49.93) --
	(221.96, 49.93) --
	(221.72, 49.93) --
	(221.48, 49.93) --
	(221.24, 49.93) --
	(221.00, 49.93) --
	(220.76, 49.93) --
	(220.52, 49.93) --
	(220.28, 49.93) --
	(220.04, 49.93) --
	(219.80, 49.93) --
	(219.56, 49.93) --
	(219.32, 49.93) --
	(219.08, 49.93) --
	(218.84, 49.93) --
	(218.60, 49.93) --
	(218.36, 49.93) --
	(218.12, 49.93) --
	(217.88, 49.94) --
	(217.64, 49.94) --
	(217.40, 49.94) --
	(217.16, 49.94) --
	(216.92, 49.94) --
	(216.68, 49.94) --
	(216.44, 49.94) --
	(216.20, 49.94) --
	(215.96, 49.95) --
	(215.72, 49.95) --
	(215.48, 49.95) --
	(215.24, 49.95) --
	(215.00, 49.95) --
	(214.76, 49.95) --
	(214.52, 49.95) --
	(214.28, 49.96) --
	(214.04, 49.96) --
	(213.80, 49.96) --
	(213.56, 49.96) --
	(213.32, 49.96) --
	(213.08, 49.96) --
	(212.84, 49.95) --
	(212.60, 49.95) --
	(212.36, 49.95) --
	(212.12, 49.95) --
	(211.88, 49.95) --
	(211.64, 49.95) --
	(211.40, 49.95) --
	(211.16, 49.94) --
	(210.92, 49.94) --
	(210.68, 49.94) --
	(210.44, 49.94) --
	(210.20, 49.94) --
	(209.96, 49.94) --
	(209.72, 49.94) --
	(209.48, 49.94) --
	(209.24, 49.93) --
	(209.00, 49.93) --
	(208.76, 49.93) --
	(208.52, 49.93) --
	(208.28, 49.93) --
	(208.04, 49.93) --
	(207.80, 49.93) --
	(207.56, 49.93) --
	(207.32, 49.93) --
	(207.08, 49.93) --
	(206.84, 49.93) --
	(206.60, 49.93) --
	(206.36, 49.93) --
	(206.12, 49.93) --
	(205.88, 49.93) --
	(205.64, 49.93) --
	(205.40, 49.93) --
	(205.16, 49.93) --
	(204.92, 49.93) --
	(204.68, 49.93) --
	(204.44, 49.93) --
	(204.20, 49.93) --
	(203.96, 49.93) --
	(203.72, 49.93) --
	(203.48, 49.93) --
	(203.24, 49.94) --
	(203.00, 49.94) --
	(202.76, 49.94) --
	(202.52, 49.94) --
	(202.28, 49.94) --
	(202.04, 49.94) --
	(201.80, 49.95) --
	(201.56, 49.95) --
	(201.32, 49.95) --
	(201.08, 49.95) --
	(200.84, 49.96) --
	(200.60, 49.96) --
	(200.36, 49.96) --
	(200.12, 49.97) --
	(199.88, 49.97) --
	(199.64, 49.97) --
	(199.40, 49.98) --
	(199.16, 49.98) --
	(198.92, 49.98) --
	(198.68, 49.98) --
	(198.44, 49.98) --
	(198.20, 49.98) --
	(197.96, 49.98) --
	(197.72, 49.98) --
	(197.48, 49.98) --
	(197.24, 49.98) --
	(197.00, 49.98) --
	(196.76, 49.98) --
	(196.52, 49.98) --
	(196.28, 49.98) --
	(196.04, 49.98) --
	(195.80, 49.98) --
	(195.56, 49.98) --
	(195.32, 49.98) --
	(195.08, 49.98) --
	(194.84, 49.98) --
	(194.60, 49.98) --
	(194.36, 49.99) --
	(194.12, 49.99) --
	(193.88, 50.00) --
	(193.64, 50.00) --
	(193.40, 50.01) --
	(193.16, 50.01) --
	(192.92, 50.02) --
	(192.68, 50.02) --
	(192.44, 50.03) --
	(192.20, 50.03) --
	(191.96, 50.04) --
	(191.72, 50.04) --
	(191.48, 50.05) --
	(191.24, 50.05) --
	(191.00, 50.05) --
	(190.76, 50.05) --
	(190.52, 50.05) --
	(190.28, 50.06) --
	(190.04, 50.06) --
	(189.80, 50.05) --
	(189.56, 50.05) --
	(189.32, 50.05) --
	(189.08, 50.05) --
	(188.84, 50.05) --
	(188.60, 50.04) --
	(188.36, 50.04) --
	(188.12, 50.04) --
	(187.88, 50.04) --
	(187.64, 50.03) --
	(187.40, 50.03) --
	(187.16, 50.03) --
	(186.92, 50.03) --
	(186.68, 50.03) --
	(186.44, 50.03) --
	(186.20, 50.03) --
	(185.96, 50.03) --
	(185.72, 50.03) --
	(185.48, 50.04) --
	(185.24, 50.04) --
	(185.00, 50.04) --
	(184.76, 50.05) --
	(184.52, 50.06) --
	(184.28, 50.06) --
	(184.04, 50.07) --
	(183.80, 50.08) --
	(183.56, 50.09) --
	(183.32, 50.10) --
	(183.08, 50.11) --
	(182.84, 50.12) --
	(182.60, 50.13) --
	(182.36, 50.14) --
	(182.12, 50.16) --
	(181.88, 50.17) --
	(181.64, 50.18) --
	(181.40, 50.19) --
	(181.16, 50.20) --
	(180.92, 50.21) --
	(180.68, 50.22) --
	(180.44, 50.23) --
	(180.20, 50.24) --
	(179.96, 50.25) --
	(179.72, 50.26) --
	(179.48, 50.26) --
	(179.24, 50.27) --
	(179.00, 50.28) --
	(178.76, 50.28) --
	(178.52, 50.29) --
	(178.28, 50.29) --
	(178.04, 50.30) --
	(177.80, 50.30) --
	(177.56, 50.31) --
	(177.32, 50.31) --
	(177.08, 50.32) --
	(176.84, 50.33) --
	(176.60, 50.33) --
	(176.36, 50.34) --
	(176.12, 50.35) --
	(175.88, 50.36) --
	(175.64, 50.37) --
	(175.40, 50.38) --
	(175.16, 50.39) --
	(174.92, 50.40) --
	(174.68, 50.41) --
	(174.44, 50.42) --
	(174.20, 50.43) --
	(173.96, 50.44) --
	(173.72, 50.45) --
	(173.48, 50.47) --
	(173.24, 50.48) --
	(173.00, 50.49) --
	(172.76, 50.50) --
	(172.52, 50.52) --
	(172.28, 50.53) --
	(172.04, 50.55) --
	(171.80, 50.57) --
	(171.56, 50.59) --
	(171.32, 50.61) --
	(171.08, 50.64) --
	(170.84, 50.66) --
	(170.60, 50.69) --
	(170.36, 50.72) --
	(170.12, 50.75) --
	(169.88, 50.78) --
	(169.64, 50.81) --
	(169.40, 50.85) --
	(169.16, 50.88) --
	(168.92, 50.91) --
	(168.68, 50.95) --
	(168.44, 50.98) --
	(168.20, 51.02) --
	(167.96, 51.05) --
	(167.72, 51.09) --
	(167.48, 51.12) --
	(167.24, 51.16) --
	(167.00, 51.20) --
	(166.76, 51.24) --
	(166.52, 51.28) --
	(166.28, 51.32) --
	(166.04, 51.36) --
	(165.80, 51.41) --
	(165.56, 51.45) --
	(165.32, 51.50) --
	(165.08, 51.55) --
	(164.84, 51.59) --
	(164.60, 51.64) --
	(164.36, 51.69) --
	(164.12, 51.73) --
	(163.88, 51.77) --
	(163.64, 51.81) --
	(163.40, 51.85) --
	(163.16, 51.89) --
	(162.92, 51.92) --
	(162.68, 51.95) --
	(162.44, 51.98) --
	(162.20, 52.01) --
	(161.96, 52.03) --
	(161.72, 52.05) --
	(161.48, 52.08) --
	(161.24, 52.10) --
	(161.00, 52.12) --
	(160.76, 52.15) --
	(160.52, 52.18) --
	(160.28, 52.21) --
	(160.04, 52.24) --
	(159.80, 52.27) --
	(159.56, 52.31) --
	(159.32, 52.36) --
	(159.08, 52.41) --
	(158.84, 52.46) --
	(158.60, 52.52) --
	(158.36, 52.59) --
	(158.12, 52.66) --
	(157.88, 52.74) --
	(157.64, 52.82) --
	(157.40, 52.91) --
	(157.16, 53.01) --
	(156.92, 53.11) --
	(156.68, 53.22) --
	(156.44, 53.33) --
	(156.20, 53.44) --
	(155.96, 53.56) --
	(155.72, 53.69) --
	(155.48, 53.81) --
	(155.24, 53.94) --
	(155.00, 54.06) --
	(154.76, 54.19) --
	(154.52, 54.32) --
	(154.28, 54.44) --
	(154.04, 54.56) --
	(153.80, 54.69) --
	(153.56, 54.80) --
	(153.32, 54.92) --
	(153.08, 55.03) --
	(152.84, 55.14) --
	(152.60, 55.24) --
	(152.36, 55.35) --
	(152.12, 55.45) --
	(151.88, 55.54) --
	(151.64, 55.64) --
	(151.40, 55.73) --
	(151.16, 55.82) --
	(150.92, 55.91) --
	(150.68, 56.00) --
	(150.44, 56.08) --
	(150.20, 56.16) --
	(149.96, 56.25) --
	(149.72, 56.33) --
	(149.48, 56.41) --
	(149.24, 56.49) --
	(149.00, 56.57) --
	(148.76, 56.65) --
	(148.52, 56.73) --
	(148.28, 56.81) --
	(148.04, 56.90) --
	(147.80, 56.98) --
	(147.56, 57.08) --
	(147.32, 57.17) --
	(147.08, 57.27) --
	(146.84, 57.38) --
	(146.60, 57.48) --
	(146.36, 57.60) --
	(146.12, 57.71) --
	(145.88, 57.84) --
	(145.64, 57.96) --
	(145.40, 58.09) --
	(145.16, 58.22) --
	(144.92, 58.35) --
	(144.68, 58.48) --
	(144.44, 58.61) --
	(144.20, 58.74) --
	(143.96, 58.87) --
	(143.72, 58.99) --
	(143.48, 59.12) --
	(143.24, 59.24) --
	(143.00, 59.35) --
	(142.76, 59.47) --
	(142.52, 59.57) --
	(142.28, 59.68) --
	(142.04, 59.78) --
	(141.80, 59.87) --
	(141.56, 59.96) --
	(141.32, 60.04) --
	(141.08, 60.12) --
	(140.84, 60.20) --
	(140.60, 60.26) --
	(140.36, 60.32) --
	(140.12, 60.37) --
	(139.88, 60.42) --
	(139.64, 60.46) --
	(139.40, 60.49) --
	(139.16, 60.52) --
	(138.92, 60.54) --
	(138.68, 60.55) --
	(138.44, 60.55) --
	(138.20, 60.54) --
	(137.96, 60.53) --
	(137.72, 60.51) --
	(137.48, 60.48) --
	(137.24, 60.45) --
	(137.00, 60.41) --
	(136.76, 60.36) --
	(136.52, 60.31) --
	(136.28, 60.25) --
	(136.04, 60.18) --
	(135.80, 60.12) --
	(135.56, 60.04) --
	(135.32, 59.96) --
	(135.08, 59.88) --
	(134.84, 59.79) --
	(134.60, 59.70) --
	(134.36, 59.60) --
	(134.12, 59.49) --
	(133.88, 59.38) --
	(133.64, 59.26) --
	(133.40, 59.14) --
	(133.16, 59.01) --
	(132.92, 58.87) --
	(132.68, 58.73) --
	(132.44, 58.59) --
	(132.20, 58.43) --
	(131.96, 58.28) --
	(131.72, 58.12) --
	(131.48, 57.96) --
	(131.24, 57.79) --
	(131.00, 57.63) --
	(130.76, 57.46) --
	(130.52, 57.30) --
	(130.28, 57.13) --
	(130.04, 56.97) --
	(129.80, 56.80) --
	(129.56, 56.64) --
	(129.32, 56.48) --
	(129.08, 56.32) --
	(128.84, 56.16) --
	(128.60, 56.00) --
	(128.36, 55.84) --
	(128.12, 55.69) --
	(127.88, 55.53) --
	(127.64, 55.37) --
	(127.40, 55.22) --
	(127.16, 55.06) --
	(126.92, 54.90) --
	(126.68, 54.75) --
	(126.44, 54.59) --
	(126.20, 54.44) --
	(125.96, 54.28) --
	(125.72, 54.13) --
	(125.48, 53.98) --
	(125.24, 53.84) --
	(125.00, 53.70) --
	(124.76, 53.56) --
	(124.52, 53.43) --
	(124.28, 53.31) --
	(124.04, 53.19) --
	(123.80, 53.07) --
	(123.56, 52.97) --
	(123.32, 52.87) --
	(123.08, 52.77) --
	(122.84, 52.68) --
	(122.60, 52.59) --
	(122.36, 52.51) --
	(122.12, 52.43) --
	(121.88, 52.36) --
	(121.64, 52.28) --
	(121.40, 52.21) --
	(121.16, 52.14) --
	(120.92, 52.06) --
	(120.68, 51.99) --
	(120.44, 51.92) --
	(120.21, 51.85) --
	(119.97, 51.78) --
	(119.73, 51.71) --
	(119.49, 51.63) --
	(119.25, 51.56) --
	(119.01, 51.49) --
	(118.77, 51.42) --
	(118.53, 51.36) --
	(118.29, 51.29) --
	(118.05, 51.23) --
	(117.81, 51.17) --
	(117.57, 51.12) --
	(117.33, 51.07) --
	(117.09, 51.02) --
	(116.85, 50.98) --
	(116.61, 50.93) --
	(116.37, 50.90) --
	(116.13, 50.86) --
	(115.89, 50.83) --
	(115.65, 50.80) --
	(115.41, 50.78) --
	(115.17, 50.75) --
	(114.93, 50.73) --
	(114.69, 50.70) --
	(114.45, 50.68) --
	(114.21, 50.66) --
	(113.97, 50.64) --
	(113.73, 50.62) --
	(113.49, 50.60) --
	(113.25, 50.58) --
	(113.01, 50.56) --
	(112.77, 50.54) --
	(112.53, 50.52) --
	(112.29, 50.50) --
	(112.05, 50.48) --
	(111.81, 50.46) --
	(111.57, 50.44) --
	(111.33, 50.41) --
	(111.09, 50.39) --
	(110.85, 50.37) --
	(110.61, 50.35) --
	(110.37, 50.33) --
	(110.13, 50.30) --
	(109.89, 50.28) --
	(109.65, 50.26) --
	(109.41, 50.24) --
	(109.17, 50.22) --
	(108.93, 50.20) --
	(108.69, 50.18) --
	(108.45, 50.17) --
	(108.21, 50.15) --
	(107.97, 50.13) --
	(107.73, 50.12) --
	(107.49, 50.11) --
	(107.25, 50.09) --
	(107.01, 50.08) --
	(106.77, 50.07) --
	(106.53, 50.06) --
	(106.29, 50.05) --
	(106.05, 50.04) --
	(105.81, 50.03) --
	(105.57, 50.02) --
	(105.33, 50.01) --
	(105.33, 49.93) --
	cycle;
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 71.72, 37.19) rectangle (235.40,211.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77, 46.90) {Pooled estimate};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77, 68.13) {Study 1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77, 89.37) {Study 2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,110.60) {Study 3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,131.84) {Study 4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,153.07) {Study 5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,174.30) {Study 6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,195.54) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97, 49.93) --
	( 71.72, 49.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97, 71.16) --
	( 71.72, 71.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97, 92.40) --
	( 71.72, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,113.63) --
	( 71.72,113.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,134.87) --
	( 71.72,134.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,156.10) --
	( 71.72,156.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,177.34) --
	( 71.72,177.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,198.57) --
	( 71.72,198.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 74.56, 34.44) --
	( 74.56, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.56, 34.44) --
	(112.56, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.55, 34.44) --
	(150.55, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.55, 34.44) --
	(188.55, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.55, 34.44) --
	(226.55, 37.19);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 74.56, 26.18) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.56, 26.18) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (150.55, 26.18) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (188.55, 26.18) {0.3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (226.55, 26.18) {0.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 70.64, 11.26) {E};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 78.12, 11.26) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 83.92, 11.26) {p};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 90.03, 11.26) {o};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 95.53, 11.26) {n};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (101.64, 11.26) {e};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (106.53, 11.26) {n};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (112.64, 11.26) {t};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (116.92, 11.26) {i};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (119.97, 11.26) {a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (125.47, 11.26) {l};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (128.52, 11.26) { };

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (134.02, 11.26) {g};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (139.52, 11.26) {r};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (143.83, 11.26) {o};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (149.33, 11.26) {w};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (157.27, 11.26) {t};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (161.55, 11.26) {h};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (167.66, 11.26) { };

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (173.16, 11.26) {r};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (177.46, 11.26) {a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (182.96, 11.26) {t};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (187.24, 11.26) {e};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (192.13, 11.26) { };

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.38] at (197.62, 11.26) {(};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (202.97, 11.26) {d};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (209.08, 11.26) {a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (214.58, 11.26) {y};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (220.38, 11.26) {s};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.77] at (224.72, 15.76) {-};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.77] at (227.29, 15.76) {1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.38] at (231.14, 11.26) {)};
\end{scope}
\begin{scope}
\path[clip] (240.90,  0.00) rectangle (481.80,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (240.90,  0.00) rectangle (481.80,216.81);
\end{scope}
\begin{scope}
\path[clip] (312.62, 31.02) rectangle (476.30,211.31);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (312.62, 31.02) rectangle (476.30,211.31);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (337.45, 31.02) --
	(337.45,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (375.96, 31.02) --
	(375.96,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (414.47, 31.02) --
	(414.47,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (452.97, 31.02) --
	(452.97,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62, 44.21) --
	(476.30, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62, 66.20) --
	(476.30, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62, 88.18) --
	(476.30, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,110.17) --
	(476.30,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,132.16) --
	(476.30,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,154.14) --
	(476.30,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,176.13) --
	(476.30,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,198.12) --
	(476.30,198.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (318.20, 31.02) --
	(318.20,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.71, 31.02) --
	(356.71,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (395.21, 31.02) --
	(395.21,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.72, 31.02) --
	(433.72,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.23, 31.02) --
	(472.23,211.31);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (395.21, 66.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (364.41, 88.18) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (364.41,110.17) circle (  2.50);
\definecolor{drawColor}{RGB}{255,165,0}
\definecolor{fillColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (356.66,198.12) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (349.00,132.16) circle (  2.50);
\definecolor{drawColor}{RGB}{255,165,0}
\definecolor{fillColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (404.72,154.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (356.69,176.13) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) --
	(395.21, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) --
	(395.21, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) --
	(395.21, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) --
	(364.41, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) --
	(364.41, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) --
	(364.41, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) --
	(364.41,110.17);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) --
	(364.41,110.17);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) --
	(364.41,110.17);
\definecolor{drawColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (393.26,198.12) --
	(393.26,198.12);

\path[draw=drawColor,line width= 0.8pt,line join=round] (393.26,198.12) --
	(320.06,198.12);

\path[draw=drawColor,line width= 0.8pt,line join=round] (320.06,198.12) --
	(320.06,198.12);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) --
	(349.00,132.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) --
	(349.00,132.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) --
	(349.00,132.16);
\definecolor{drawColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (468.86,154.14) --
	(468.86,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (468.86,154.14) --
	(340.83,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (340.83,154.14) --
	(340.83,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.04,176.13) --
	(364.04,176.13);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.04,176.13) --
	(349.40,176.13);

\path[draw=drawColor,line width= 0.8pt,line join=round] (349.40,176.13) --
	(349.40,176.13);
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (327.85, 44.21) --
	(328.07, 44.21) --
	(328.28, 44.21) --
	(328.50, 44.21) --
	(328.72, 44.21) --
	(328.93, 44.21) --
	(329.15, 44.21) --
	(329.36, 44.21) --
	(329.58, 44.21) --
	(329.79, 44.21) --
	(330.01, 44.21) --
	(330.23, 44.21) --
	(330.44, 44.21) --
	(330.66, 44.21) --
	(330.87, 44.21) --
	(331.09, 44.21) --
	(331.30, 44.21) --
	(331.52, 44.21) --
	(331.74, 44.21) --
	(331.95, 44.21) --
	(332.17, 44.21) --
	(332.38, 44.21) --
	(332.60, 44.21) --
	(332.81, 44.21) --
	(333.03, 44.21) --
	(333.25, 44.21) --
	(333.46, 44.21) --
	(333.68, 44.21) --
	(333.89, 44.21) --
	(334.11, 44.21) --
	(334.32, 44.21) --
	(334.54, 44.21) --
	(334.76, 44.21) --
	(334.97, 44.21) --
	(335.19, 44.21) --
	(335.40, 44.21) --
	(335.62, 44.21) --
	(335.83, 44.21) --
	(336.05, 44.21) --
	(336.27, 44.21) --
	(336.48, 44.21) --
	(336.70, 44.21) --
	(336.91, 44.21) --
	(337.13, 44.21) --
	(337.34, 44.21) --
	(337.56, 44.21) --
	(337.78, 44.21) --
	(337.99, 44.21) --
	(338.21, 44.21) --
	(338.42, 44.21) --
	(338.64, 44.21) --
	(338.86, 44.21) --
	(339.07, 44.21) --
	(339.29, 44.21) --
	(339.50, 44.21) --
	(339.72, 44.21) --
	(339.93, 44.21) --
	(340.15, 44.21) --
	(340.37, 44.21) --
	(340.58, 44.21) --
	(340.80, 44.21) --
	(341.01, 44.21) --
	(341.23, 44.21) --
	(341.44, 44.21) --
	(341.66, 44.21) --
	(341.88, 44.21) --
	(342.09, 44.21) --
	(342.31, 44.21) --
	(342.52, 44.21) --
	(342.74, 44.21) --
	(342.95, 44.21) --
	(343.17, 44.21) --
	(343.39, 44.21) --
	(343.60, 44.21) --
	(343.82, 44.21) --
	(344.03, 44.21) --
	(344.25, 44.21) --
	(344.46, 44.21) --
	(344.68, 44.21) --
	(344.90, 44.21) --
	(345.11, 44.21) --
	(345.33, 44.21) --
	(345.54, 44.21) --
	(345.76, 44.21) --
	(345.97, 44.21) --
	(346.19, 44.21) --
	(346.41, 44.21) --
	(346.62, 44.21) --
	(346.84, 44.21) --
	(347.05, 44.21) --
	(347.27, 44.21) --
	(347.48, 44.21) --
	(347.70, 44.21) --
	(347.92, 44.21) --
	(348.13, 44.21) --
	(348.35, 44.21) --
	(348.56, 44.21) --
	(348.78, 44.21) --
	(348.99, 44.21) --
	(349.21, 44.21) --
	(349.43, 44.21) --
	(349.64, 44.21) --
	(349.86, 44.21) --
	(350.07, 44.21) --
	(350.29, 44.21) --
	(350.50, 44.21) --
	(350.72, 44.21) --
	(350.94, 44.21) --
	(351.15, 44.21) --
	(351.37, 44.21) --
	(351.58, 44.21) --
	(351.80, 44.21) --
	(352.02, 44.21) --
	(352.23, 44.21) --
	(352.45, 44.21) --
	(352.66, 44.21) --
	(352.88, 44.21) --
	(353.09, 44.21) --
	(353.31, 44.21) --
	(353.53, 44.21) --
	(353.74, 44.21) --
	(353.96, 44.21) --
	(354.17, 44.21) --
	(354.39, 44.21) --
	(354.60, 44.21) --
	(354.82, 44.21) --
	(355.04, 44.21) --
	(355.25, 44.21) --
	(355.47, 44.21) --
	(355.68, 44.21) --
	(355.90, 44.21) --
	(356.11, 44.21) --
	(356.33, 44.21) --
	(356.55, 44.21) --
	(356.76, 44.21) --
	(356.98, 44.21) --
	(357.19, 44.21) --
	(357.41, 44.21) --
	(357.62, 44.21) --
	(357.84, 44.21) --
	(358.06, 44.21) --
	(358.27, 44.21) --
	(358.49, 44.21) --
	(358.70, 44.21) --
	(358.92, 44.21) --
	(359.13, 44.21) --
	(359.35, 44.21) --
	(359.57, 44.21) --
	(359.78, 44.21) --
	(360.00, 44.21) --
	(360.21, 44.21) --
	(360.43, 44.21) --
	(360.64, 44.21) --
	(360.86, 44.21) --
	(361.08, 44.21) --
	(361.29, 44.21) --
	(361.51, 44.21) --
	(361.72, 44.21) --
	(361.94, 44.21) --
	(362.15, 44.21) --
	(362.37, 44.21) --
	(362.59, 44.21) --
	(362.80, 44.21) --
	(363.02, 44.21) --
	(363.23, 44.21) --
	(363.45, 44.21) --
	(363.66, 44.21) --
	(363.88, 44.21) --
	(364.10, 44.21) --
	(364.31, 44.21) --
	(364.53, 44.21) --
	(364.74, 44.21) --
	(364.96, 44.21) --
	(365.18, 44.21) --
	(365.39, 44.21) --
	(365.61, 44.21) --
	(365.82, 44.21) --
	(366.04, 44.21) --
	(366.25, 44.21) --
	(366.47, 44.21) --
	(366.69, 44.21) --
	(366.90, 44.21) --
	(367.12, 44.21) --
	(367.33, 44.21) --
	(367.55, 44.21) --
	(367.76, 44.21) --
	(367.98, 44.21) --
	(368.20, 44.21) --
	(368.41, 44.21) --
	(368.63, 44.21) --
	(368.84, 44.21) --
	(369.06, 44.21) --
	(369.27, 44.21) --
	(369.49, 44.21) --
	(369.71, 44.21) --
	(369.92, 44.21) --
	(370.14, 44.21) --
	(370.35, 44.21) --
	(370.57, 44.21) --
	(370.78, 44.21) --
	(371.00, 44.21) --
	(371.22, 44.21) --
	(371.43, 44.21) --
	(371.65, 44.21) --
	(371.86, 44.21) --
	(372.08, 44.21) --
	(372.29, 44.21) --
	(372.51, 44.21) --
	(372.73, 44.21) --
	(372.94, 44.21) --
	(373.16, 44.21) --
	(373.37, 44.21) --
	(373.59, 44.21) --
	(373.80, 44.21) --
	(374.02, 44.21) --
	(374.24, 44.21) --
	(374.45, 44.21) --
	(374.67, 44.21) --
	(374.88, 44.21) --
	(375.10, 44.21) --
	(375.31, 44.21) --
	(375.53, 44.21) --
	(375.75, 44.21) --
	(375.96, 44.21) --
	(376.18, 44.21) --
	(376.39, 44.21) --
	(376.61, 44.21) --
	(376.82, 44.21) --
	(377.04, 44.21) --
	(377.26, 44.21) --
	(377.47, 44.21) --
	(377.69, 44.21) --
	(377.90, 44.21) --
	(378.12, 44.21) --
	(378.33, 44.21) --
	(378.55, 44.21) --
	(378.77, 44.21) --
	(378.98, 44.21) --
	(379.20, 44.21) --
	(379.41, 44.21) --
	(379.63, 44.21) --
	(379.85, 44.21) --
	(380.06, 44.21) --
	(380.28, 44.21) --
	(380.49, 44.21) --
	(380.71, 44.21) --
	(380.92, 44.21) --
	(381.14, 44.21) --
	(381.36, 44.21) --
	(381.57, 44.21) --
	(381.79, 44.21) --
	(382.00, 44.21) --
	(382.22, 44.21) --
	(382.43, 44.21) --
	(382.65, 44.21) --
	(382.87, 44.21) --
	(383.08, 44.21) --
	(383.30, 44.21) --
	(383.51, 44.21) --
	(383.73, 44.21) --
	(383.94, 44.21) --
	(384.16, 44.21) --
	(384.38, 44.21) --
	(384.59, 44.21) --
	(384.81, 44.21) --
	(385.02, 44.21) --
	(385.24, 44.21) --
	(385.45, 44.21) --
	(385.67, 44.21) --
	(385.89, 44.21) --
	(386.10, 44.21) --
	(386.32, 44.21) --
	(386.53, 44.21) --
	(386.75, 44.21) --
	(386.96, 44.21) --
	(387.18, 44.21) --
	(387.40, 44.21) --
	(387.61, 44.21) --
	(387.83, 44.21) --
	(388.04, 44.21) --
	(388.26, 44.21) --
	(388.47, 44.21) --
	(388.69, 44.21) --
	(388.91, 44.21) --
	(389.12, 44.21) --
	(389.34, 44.21) --
	(389.55, 44.21) --
	(389.77, 44.21) --
	(389.98, 44.21) --
	(390.20, 44.21) --
	(390.42, 44.21) --
	(390.63, 44.21) --
	(390.85, 44.21) --
	(391.06, 44.21) --
	(391.28, 44.21) --
	(391.49, 44.21) --
	(391.71, 44.21) --
	(391.93, 44.21) --
	(392.14, 44.21) --
	(392.36, 44.21) --
	(392.57, 44.21) --
	(392.79, 44.21) --
	(393.01, 44.21) --
	(393.22, 44.21) --
	(393.44, 44.21) --
	(393.65, 44.21) --
	(393.87, 44.21) --
	(394.08, 44.21) --
	(394.30, 44.21) --
	(394.52, 44.21) --
	(394.73, 44.21) --
	(394.95, 44.21) --
	(395.16, 44.21) --
	(395.38, 44.21) --
	(395.59, 44.21) --
	(395.81, 44.21) --
	(396.03, 44.21) --
	(396.24, 44.21) --
	(396.46, 44.21) --
	(396.67, 44.21) --
	(396.89, 44.21) --
	(397.10, 44.21) --
	(397.32, 44.21) --
	(397.54, 44.21) --
	(397.75, 44.21) --
	(397.97, 44.21) --
	(398.18, 44.21) --
	(398.40, 44.21) --
	(398.61, 44.21) --
	(398.83, 44.21) --
	(399.05, 44.21) --
	(399.26, 44.21) --
	(399.48, 44.21) --
	(399.69, 44.21) --
	(399.91, 44.21) --
	(400.12, 44.21) --
	(400.34, 44.21) --
	(400.56, 44.21) --
	(400.77, 44.21) --
	(400.99, 44.21) --
	(401.20, 44.21) --
	(401.42, 44.21) --
	(401.63, 44.21) --
	(401.85, 44.21) --
	(402.07, 44.21) --
	(402.28, 44.21) --
	(402.50, 44.21) --
	(402.71, 44.21) --
	(402.93, 44.21) --
	(403.14, 44.21) --
	(403.36, 44.21) --
	(403.58, 44.21) --
	(403.79, 44.21) --
	(404.01, 44.21) --
	(404.22, 44.21) --
	(404.44, 44.21) --
	(404.65, 44.21) --
	(404.87, 44.21) --
	(405.09, 44.21) --
	(405.30, 44.21) --
	(405.52, 44.21) --
	(405.73, 44.21) --
	(405.95, 44.21) --
	(406.16, 44.21) --
	(406.38, 44.21) --
	(406.60, 44.21) --
	(406.81, 44.21) --
	(407.03, 44.21) --
	(407.24, 44.21) --
	(407.46, 44.21) --
	(407.68, 44.21) --
	(407.89, 44.21) --
	(408.11, 44.21) --
	(408.32, 44.21) --
	(408.54, 44.21) --
	(408.75, 44.21) --
	(408.97, 44.21) --
	(409.19, 44.21) --
	(409.40, 44.21) --
	(409.62, 44.21) --
	(409.83, 44.21) --
	(410.05, 44.21) --
	(410.26, 44.21) --
	(410.48, 44.21) --
	(410.70, 44.21) --
	(410.91, 44.21) --
	(411.13, 44.21) --
	(411.34, 44.21) --
	(411.56, 44.21) --
	(411.77, 44.21) --
	(411.99, 44.21) --
	(412.21, 44.21) --
	(412.42, 44.21) --
	(412.64, 44.21) --
	(412.85, 44.21) --
	(413.07, 44.21) --
	(413.28, 44.21) --
	(413.50, 44.21) --
	(413.72, 44.21) --
	(413.93, 44.21) --
	(414.15, 44.21) --
	(414.36, 44.21) --
	(414.58, 44.21) --
	(414.79, 44.21) --
	(415.01, 44.21) --
	(415.23, 44.21) --
	(415.44, 44.21) --
	(415.66, 44.21) --
	(415.87, 44.21) --
	(416.09, 44.21) --
	(416.30, 44.21) --
	(416.52, 44.21) --
	(416.74, 44.21) --
	(416.95, 44.21) --
	(417.17, 44.21) --
	(417.38, 44.21) --
	(417.60, 44.21) --
	(417.81, 44.21) --
	(418.03, 44.21) --
	(418.25, 44.21) --
	(418.46, 44.21) --
	(418.68, 44.21) --
	(418.89, 44.21) --
	(419.11, 44.21) --
	(419.32, 44.21) --
	(419.54, 44.21) --
	(419.76, 44.21) --
	(419.97, 44.21) --
	(420.19, 44.21) --
	(420.40, 44.21) --
	(420.62, 44.21) --
	(420.84, 44.21) --
	(421.05, 44.21) --
	(421.27, 44.21) --
	(421.48, 44.21) --
	(421.70, 44.21) --
	(421.91, 44.21) --
	(422.13, 44.21) --
	(422.35, 44.21) --
	(422.56, 44.21) --
	(422.78, 44.21) --
	(422.99, 44.21) --
	(423.21, 44.21) --
	(423.42, 44.21) --
	(423.64, 44.21) --
	(423.86, 44.21) --
	(424.07, 44.21) --
	(424.29, 44.21) --
	(424.50, 44.21) --
	(424.72, 44.21) --
	(424.93, 44.21) --
	(425.15, 44.21) --
	(425.37, 44.21) --
	(425.58, 44.21) --
	(425.80, 44.21) --
	(426.01, 44.21) --
	(426.23, 44.21) --
	(426.44, 44.21) --
	(426.66, 44.21) --
	(426.88, 44.21) --
	(427.09, 44.21) --
	(427.31, 44.21) --
	(427.52, 44.21) --
	(427.74, 44.21) --
	(427.95, 44.21) --
	(428.17, 44.21) --
	(428.39, 44.21) --
	(428.60, 44.21) --
	(428.82, 44.21) --
	(429.03, 44.21) --
	(429.25, 44.21) --
	(429.46, 44.21) --
	(429.68, 44.21) --
	(429.90, 44.21) --
	(430.11, 44.21) --
	(430.33, 44.21) --
	(430.54, 44.21) --
	(430.76, 44.21) --
	(430.97, 44.21) --
	(431.19, 44.21) --
	(431.41, 44.21) --
	(431.62, 44.21) --
	(431.84, 44.21) --
	(432.05, 44.21) --
	(432.27, 44.21) --
	(432.48, 44.21) --
	(432.70, 44.21) --
	(432.92, 44.21) --
	(433.13, 44.21) --
	(433.35, 44.21) --
	(433.56, 44.21) --
	(433.78, 44.21) --
	(433.99, 44.21) --
	(434.21, 44.21) --
	(434.43, 44.21) --
	(434.64, 44.21) --
	(434.86, 44.21) --
	(435.07, 44.21) --
	(435.29, 44.21) --
	(435.51, 44.21) --
	(435.72, 44.21) --
	(435.94, 44.21) --
	(436.15, 44.21) --
	(436.37, 44.21) --
	(436.58, 44.21) --
	(436.80, 44.21) --
	(437.02, 44.21) --
	(437.23, 44.21) --
	(437.45, 44.21) --
	(437.66, 44.21) --
	(437.88, 44.21) --
	(438.09, 44.21) --
	(438.09, 44.24) --
	(437.88, 44.24) --
	(437.66, 44.24) --
	(437.45, 44.24) --
	(437.23, 44.24) --
	(437.02, 44.23) --
	(436.80, 44.23) --
	(436.58, 44.23) --
	(436.37, 44.23) --
	(436.15, 44.22) --
	(435.94, 44.22) --
	(435.72, 44.22) --
	(435.51, 44.22) --
	(435.29, 44.22) --
	(435.07, 44.22) --
	(434.86, 44.22) --
	(434.64, 44.21) --
	(434.43, 44.21) --
	(434.21, 44.21) --
	(433.99, 44.21) --
	(433.78, 44.21) --
	(433.56, 44.21) --
	(433.35, 44.21) --
	(433.13, 44.21) --
	(432.92, 44.21) --
	(432.70, 44.21) --
	(432.48, 44.21) --
	(432.27, 44.22) --
	(432.05, 44.22) --
	(431.84, 44.22) --
	(431.62, 44.22) --
	(431.41, 44.22) --
	(431.19, 44.22) --
	(430.97, 44.23) --
	(430.76, 44.23) --
	(430.54, 44.23) --
	(430.33, 44.23) --
	(430.11, 44.23) --
	(429.90, 44.24) --
	(429.68, 44.24) --
	(429.46, 44.24) --
	(429.25, 44.24) --
	(429.03, 44.24) --
	(428.82, 44.24) --
	(428.60, 44.24) --
	(428.39, 44.24) --
	(428.17, 44.23) --
	(427.95, 44.23) --
	(427.74, 44.23) --
	(427.52, 44.23) --
	(427.31, 44.23) --
	(427.09, 44.22) --
	(426.88, 44.22) --
	(426.66, 44.22) --
	(426.44, 44.22) --
	(426.23, 44.22) --
	(426.01, 44.22) --
	(425.80, 44.21) --
	(425.58, 44.21) --
	(425.37, 44.21) --
	(425.15, 44.21) --
	(424.93, 44.21) --
	(424.72, 44.21) --
	(424.50, 44.21) --
	(424.29, 44.21) --
	(424.07, 44.21) --
	(423.86, 44.21) --
	(423.64, 44.21) --
	(423.42, 44.21) --
	(423.21, 44.21) --
	(422.99, 44.21) --
	(422.78, 44.21) --
	(422.56, 44.21) --
	(422.35, 44.21) --
	(422.13, 44.21) --
	(421.91, 44.21) --
	(421.70, 44.21) --
	(421.48, 44.21) --
	(421.27, 44.21) --
	(421.05, 44.21) --
	(420.84, 44.21) --
	(420.62, 44.21) --
	(420.40, 44.21) --
	(420.19, 44.21) --
	(419.97, 44.21) --
	(419.76, 44.21) --
	(419.54, 44.21) --
	(419.32, 44.21) --
	(419.11, 44.21) --
	(418.89, 44.21) --
	(418.68, 44.21) --
	(418.46, 44.21) --
	(418.25, 44.21) --
	(418.03, 44.21) --
	(417.81, 44.21) --
	(417.60, 44.21) --
	(417.38, 44.21) --
	(417.17, 44.21) --
	(416.95, 44.21) --
	(416.74, 44.21) --
	(416.52, 44.21) --
	(416.30, 44.21) --
	(416.09, 44.21) --
	(415.87, 44.21) --
	(415.66, 44.21) --
	(415.44, 44.21) --
	(415.23, 44.21) --
	(415.01, 44.21) --
	(414.79, 44.21) --
	(414.58, 44.21) --
	(414.36, 44.21) --
	(414.15, 44.21) --
	(413.93, 44.21) --
	(413.72, 44.21) --
	(413.50, 44.21) --
	(413.28, 44.21) --
	(413.07, 44.21) --
	(412.85, 44.21) --
	(412.64, 44.21) --
	(412.42, 44.22) --
	(412.21, 44.22) --
	(411.99, 44.22) --
	(411.77, 44.22) --
	(411.56, 44.22) --
	(411.34, 44.22) --
	(411.13, 44.23) --
	(410.91, 44.23) --
	(410.70, 44.23) --
	(410.48, 44.24) --
	(410.26, 44.24) --
	(410.05, 44.24) --
	(409.83, 44.25) --
	(409.62, 44.25) --
	(409.40, 44.26) --
	(409.19, 44.26) --
	(408.97, 44.26) --
	(408.75, 44.27) --
	(408.54, 44.27) --
	(408.32, 44.27) --
	(408.11, 44.28) --
	(407.89, 44.28) --
	(407.68, 44.28) --
	(407.46, 44.29) --
	(407.24, 44.29) --
	(407.03, 44.29) --
	(406.81, 44.29) --
	(406.60, 44.30) --
	(406.38, 44.30) --
	(406.16, 44.30) --
	(405.95, 44.31) --
	(405.73, 44.31) --
	(405.52, 44.32) --
	(405.30, 44.32) --
	(405.09, 44.33) --
	(404.87, 44.34) --
	(404.65, 44.34) --
	(404.44, 44.35) --
	(404.22, 44.36) --
	(404.01, 44.36) --
	(403.79, 44.37) --
	(403.58, 44.38) --
	(403.36, 44.38) --
	(403.14, 44.39) --
	(402.93, 44.39) --
	(402.71, 44.39) --
	(402.50, 44.39) --
	(402.28, 44.39) --
	(402.07, 44.38) --
	(401.85, 44.38) --
	(401.63, 44.38) --
	(401.42, 44.37) --
	(401.20, 44.37) --
	(400.99, 44.37) --
	(400.77, 44.36) --
	(400.56, 44.36) --
	(400.34, 44.36) --
	(400.12, 44.36) --
	(399.91, 44.36) --
	(399.69, 44.36) --
	(399.48, 44.37) --
	(399.26, 44.37) --
	(399.05, 44.38) --
	(398.83, 44.38) --
	(398.61, 44.39) --
	(398.40, 44.40) --
	(398.18, 44.41) --
	(397.97, 44.42) --
	(397.75, 44.43) --
	(397.54, 44.43) --
	(397.32, 44.44) --
	(397.10, 44.45) --
	(396.89, 44.46) --
	(396.67, 44.47) --
	(396.46, 44.48) --
	(396.24, 44.48) --
	(396.03, 44.49) --
	(395.81, 44.50) --
	(395.59, 44.50) --
	(395.38, 44.51) --
	(395.16, 44.51) --
	(394.95, 44.52) --
	(394.73, 44.52) --
	(394.52, 44.52) --
	(394.30, 44.53) --
	(394.08, 44.53) --
	(393.87, 44.53) --
	(393.65, 44.54) --
	(393.44, 44.54) --
	(393.22, 44.55) --
	(393.01, 44.56) --
	(392.79, 44.57) --
	(392.57, 44.58) --
	(392.36, 44.59) --
	(392.14, 44.60) --
	(391.93, 44.62) --
	(391.71, 44.64) --
	(391.49, 44.66) --
	(391.28, 44.68) --
	(391.06, 44.70) --
	(390.85, 44.72) --
	(390.63, 44.74) --
	(390.42, 44.76) --
	(390.20, 44.78) --
	(389.98, 44.80) --
	(389.77, 44.82) --
	(389.55, 44.84) --
	(389.34, 44.86) --
	(389.12, 44.88) --
	(388.91, 44.90) --
	(388.69, 44.92) --
	(388.47, 44.94) --
	(388.26, 44.96) --
	(388.04, 44.99) --
	(387.83, 45.02) --
	(387.61, 45.05) --
	(387.40, 45.09) --
	(387.18, 45.12) --
	(386.96, 45.16) --
	(386.75, 45.21) --
	(386.53, 45.25) --
	(386.32, 45.30) --
	(386.10, 45.35) --
	(385.89, 45.40) --
	(385.67, 45.45) --
	(385.45, 45.50) --
	(385.24, 45.55) --
	(385.02, 45.60) --
	(384.81, 45.65) --
	(384.59, 45.70) --
	(384.38, 45.75) --
	(384.16, 45.80) --
	(383.94, 45.85) --
	(383.73, 45.90) --
	(383.51, 45.95) --
	(383.30, 46.01) --
	(383.08, 46.06) --
	(382.87, 46.12) --
	(382.65, 46.18) --
	(382.43, 46.24) --
	(382.22, 46.30) --
	(382.00, 46.36) --
	(381.79, 46.42) --
	(381.57, 46.47) --
	(381.36, 46.53) --
	(381.14, 46.59) --
	(380.92, 46.65) --
	(380.71, 46.70) --
	(380.49, 46.76) --
	(380.28, 46.82) --
	(380.06, 46.87) --
	(379.85, 46.94) --
	(379.63, 47.01) --
	(379.41, 47.08) --
	(379.20, 47.16) --
	(378.98, 47.25) --
	(378.77, 47.35) --
	(378.55, 47.46) --
	(378.33, 47.58) --
	(378.12, 47.71) --
	(377.90, 47.85) --
	(377.69, 48.00) --
	(377.47, 48.16) --
	(377.26, 48.33) --
	(377.04, 48.50) --
	(376.82, 48.68) --
	(376.61, 48.85) --
	(376.39, 49.03) --
	(376.18, 49.21) --
	(375.96, 49.38) --
	(375.75, 49.55) --
	(375.53, 49.72) --
	(375.31, 49.88) --
	(375.10, 50.04) --
	(374.88, 50.20) --
	(374.67, 50.35) --
	(374.45, 50.51) --
	(374.24, 50.66) --
	(374.02, 50.82) --
	(373.80, 50.98) --
	(373.59, 51.14) --
	(373.37, 51.30) --
	(373.16, 51.47) --
	(372.94, 51.63) --
	(372.73, 51.79) --
	(372.51, 51.95) --
	(372.29, 52.10) --
	(372.08, 52.25) --
	(371.86, 52.40) --
	(371.65, 52.54) --
	(371.43, 52.68) --
	(371.22, 52.81) --
	(371.00, 52.94) --
	(370.78, 53.06) --
	(370.57, 53.19) --
	(370.35, 53.31) --
	(370.14, 53.43) --
	(369.92, 53.55) --
	(369.71, 53.66) --
	(369.49, 53.78) --
	(369.27, 53.89) --
	(369.06, 54.01) --
	(368.84, 54.12) --
	(368.63, 54.22) --
	(368.41, 54.33) --
	(368.20, 54.43) --
	(367.98, 54.53) --
	(367.76, 54.62) --
	(367.55, 54.71) --
	(367.33, 54.79) --
	(367.12, 54.87) --
	(366.90, 54.95) --
	(366.69, 55.01) --
	(366.47, 55.07) --
	(366.25, 55.11) --
	(366.04, 55.15) --
	(365.82, 55.18) --
	(365.61, 55.20) --
	(365.39, 55.20) --
	(365.18, 55.20) --
	(364.96, 55.18) --
	(364.74, 55.16) --
	(364.53, 55.11) --
	(364.31, 55.06) --
	(364.10, 54.99) --
	(363.88, 54.91) --
	(363.66, 54.82) --
	(363.45, 54.72) --
	(363.23, 54.60) --
	(363.02, 54.48) --
	(362.80, 54.34) --
	(362.59, 54.19) --
	(362.37, 54.04) --
	(362.15, 53.87) --
	(361.94, 53.71) --
	(361.72, 53.53) --
	(361.51, 53.35) --
	(361.29, 53.17) --
	(361.08, 52.99) --
	(360.86, 52.80) --
	(360.64, 52.62) --
	(360.43, 52.43) --
	(360.21, 52.25) --
	(360.00, 52.06) --
	(359.78, 51.88) --
	(359.57, 51.70) --
	(359.35, 51.52) --
	(359.13, 51.35) --
	(358.92, 51.17) --
	(358.70, 50.99) --
	(358.49, 50.81) --
	(358.27, 50.63) --
	(358.06, 50.44) --
	(357.84, 50.26) --
	(357.62, 50.06) --
	(357.41, 49.87) --
	(357.19, 49.67) --
	(356.98, 49.46) --
	(356.76, 49.26) --
	(356.55, 49.05) --
	(356.33, 48.85) --
	(356.11, 48.65) --
	(355.90, 48.46) --
	(355.68, 48.27) --
	(355.47, 48.10) --
	(355.25, 47.93) --
	(355.04, 47.77) --
	(354.82, 47.63) --
	(354.60, 47.49) --
	(354.39, 47.37) --
	(354.17, 47.27) --
	(353.96, 47.17) --
	(353.74, 47.08) --
	(353.53, 47.00) --
	(353.31, 46.92) --
	(353.09, 46.85) --
	(352.88, 46.78) --
	(352.66, 46.72) --
	(352.45, 46.65) --
	(352.23, 46.58) --
	(352.02, 46.51) --
	(351.80, 46.44) --
	(351.58, 46.37) --
	(351.37, 46.30) --
	(351.15, 46.23) --
	(350.94, 46.15) --
	(350.72, 46.08) --
	(350.50, 46.01) --
	(350.29, 45.93) --
	(350.07, 45.86) --
	(349.86, 45.79) --
	(349.64, 45.72) --
	(349.43, 45.66) --
	(349.21, 45.59) --
	(348.99, 45.53) --
	(348.78, 45.47) --
	(348.56, 45.41) --
	(348.35, 45.35) --
	(348.13, 45.30) --
	(347.92, 45.25) --
	(347.70, 45.21) --
	(347.48, 45.16) --
	(347.27, 45.12) --
	(347.05, 45.09) --
	(346.84, 45.05) --
	(346.62, 45.02) --
	(346.41, 44.99) --
	(346.19, 44.96) --
	(345.97, 44.93) --
	(345.76, 44.91) --
	(345.54, 44.88) --
	(345.33, 44.85) --
	(345.11, 44.83) --
	(344.90, 44.80) --
	(344.68, 44.77) --
	(344.46, 44.74) --
	(344.25, 44.71) --
	(344.03, 44.68) --
	(343.82, 44.64) --
	(343.60, 44.61) --
	(343.39, 44.58) --
	(343.17, 44.54) --
	(342.95, 44.51) --
	(342.74, 44.48) --
	(342.52, 44.45) --
	(342.31, 44.43) --
	(342.09, 44.41) --
	(341.88, 44.39) --
	(341.66, 44.37) --
	(341.44, 44.36) --
	(341.23, 44.35) --
	(341.01, 44.34) --
	(340.80, 44.34) --
	(340.58, 44.33) --
	(340.37, 44.33) --
	(340.15, 44.34) --
	(339.93, 44.34) --
	(339.72, 44.34) --
	(339.50, 44.34) --
	(339.29, 44.35) --
	(339.07, 44.35) --
	(338.86, 44.36) --
	(338.64, 44.36) --
	(338.42, 44.36) --
	(338.21, 44.37) --
	(337.99, 44.37) --
	(337.78, 44.37) --
	(337.56, 44.37) --
	(337.34, 44.38) --
	(337.13, 44.38) --
	(336.91, 44.38) --
	(336.70, 44.38) --
	(336.48, 44.38) --
	(336.27, 44.37) --
	(336.05, 44.37) --
	(335.83, 44.37) --
	(335.62, 44.36) --
	(335.40, 44.36) --
	(335.19, 44.35) --
	(334.97, 44.35) --
	(334.76, 44.34) --
	(334.54, 44.33) --
	(334.32, 44.33) --
	(334.11, 44.33) --
	(333.89, 44.32) --
	(333.68, 44.32) --
	(333.46, 44.31) --
	(333.25, 44.31) --
	(333.03, 44.31) --
	(332.81, 44.31) --
	(332.60, 44.30) --
	(332.38, 44.30) --
	(332.17, 44.30) --
	(331.95, 44.29) --
	(331.74, 44.29) --
	(331.52, 44.28) --
	(331.30, 44.28) --
	(331.09, 44.27) --
	(330.87, 44.27) --
	(330.66, 44.27) --
	(330.44, 44.26) --
	(330.23, 44.26) --
	(330.01, 44.26) --
	(329.79, 44.25) --
	(329.58, 44.25) --
	(329.36, 44.25) --
	(329.15, 44.25) --
	(328.93, 44.25) --
	(328.72, 44.25) --
	(328.50, 44.24) --
	(328.28, 44.24) --
	(328.07, 44.24) --
	(327.85, 44.24) --
	(327.85, 44.21) --
	cycle;
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.62, 31.02) rectangle (476.30,211.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67, 41.18) {Pooled estimate};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67, 63.17) {Study 1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67, 85.15) {Study 2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,107.14) {Study 3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,129.13) {Study 4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,151.11) {Study 5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,173.10) {Study 6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,195.09) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87, 44.21) --
	(312.62, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87, 66.20) --
	(312.62, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87, 88.18) --
	(312.62, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,110.17) --
	(312.62,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,132.16) --
	(312.62,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,154.14) --
	(312.62,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,176.13) --
	(312.62,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,198.12) --
	(312.62,198.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (318.20, 28.27) --
	(318.20, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.71, 28.27) --
	(356.71, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (395.21, 28.27) --
	(395.21, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.72, 28.27) --
	(433.72, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.23, 28.27) --
	(472.23, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.20, 20.01) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (356.71, 20.01) {8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (395.21, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (433.72, 20.01) {12};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (472.23, 20.01) {14};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (394.46,  7.82) {Mean generation interval (days)};
\end{scope}
\begin{scope}
\path[clip] (481.80,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (481.80,  0.00) rectangle (722.70,216.81);
\end{scope}
\begin{scope}
\path[clip] (553.52, 31.02) rectangle (717.20,211.31);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (553.52, 31.02) rectangle (717.20,211.31);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (579.31, 31.02) --
	(579.31,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (616.03, 31.02) --
	(616.03,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (652.74, 31.02) --
	(652.74,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (689.45, 31.02) --
	(689.45,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52, 44.21) --
	(717.20, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52, 66.20) --
	(717.20, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52, 88.18) --
	(717.20, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,110.17) --
	(717.20,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,132.16) --
	(717.20,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,154.14) --
	(717.20,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,176.13) --
	(717.20,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,198.12) --
	(717.20,198.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (560.96, 31.02) --
	(560.96,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.67, 31.02) --
	(597.67,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.38, 31.02) --
	(634.38,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (671.09, 31.02) --
	(671.09,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (707.80, 31.02) --
	(707.80,211.31);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (634.38, 66.20) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (597.67, 92.07) --
	(601.03, 86.24) --
	(594.31, 86.24) --
	(597.67, 92.07);

\path[fill=fillColor] (575.98,110.17) circle (  2.50);

\path[fill=fillColor] (560.96,198.12) circle (  2.50);

\path[fill=fillColor] (597.67,132.16) circle (  2.50);

\path[fill=fillColor] (597.67,154.14) circle (  2.50);

\path[fill=fillColor] (575.64,176.13) circle (  2.50);

\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) --
	(634.38, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) --
	(634.38, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) --
	(634.38, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) --
	(597.67, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) --
	(597.67, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) --
	(597.67, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) --
	(575.98,110.17);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) --
	(575.98,110.17);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) --
	(575.98,110.17);

\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) --
	(560.96,198.12);

\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) --
	(560.96,198.12);

\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) --
	(560.96,198.12);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) --
	(597.67,132.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) --
	(597.67,132.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) --
	(597.67,132.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) --
	(597.67,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) --
	(597.67,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) --
	(597.67,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) --
	(575.64,176.13);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) --
	(575.64,176.13);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) --
	(575.64,176.13);
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (571.29, 44.21) --
	(571.56, 44.21) --
	(571.83, 44.21) --
	(572.10, 44.21) --
	(572.37, 44.21) --
	(572.65, 44.21) --
	(572.92, 44.21) --
	(573.19, 44.21) --
	(573.46, 44.21) --
	(573.73, 44.21) --
	(574.00, 44.21) --
	(574.27, 44.21) --
	(574.54, 44.21) --
	(574.81, 44.21) --
	(575.08, 44.21) --
	(575.36, 44.21) --
	(575.63, 44.21) --
	(575.90, 44.21) --
	(576.17, 44.21) --
	(576.44, 44.21) --
	(576.71, 44.21) --
	(576.98, 44.21) --
	(577.25, 44.21) --
	(577.52, 44.21) --
	(577.79, 44.21) --
	(578.07, 44.21) --
	(578.34, 44.21) --
	(578.61, 44.21) --
	(578.88, 44.21) --
	(579.15, 44.21) --
	(579.42, 44.21) --
	(579.69, 44.21) --
	(579.96, 44.21) --
	(580.23, 44.21) --
	(580.50, 44.21) --
	(580.77, 44.21) --
	(581.05, 44.21) --
	(581.32, 44.21) --
	(581.59, 44.21) --
	(581.86, 44.21) --
	(582.13, 44.21) --
	(582.40, 44.21) --
	(582.67, 44.21) --
	(582.94, 44.21) --
	(583.21, 44.21) --
	(583.48, 44.21) --
	(583.76, 44.21) --
	(584.03, 44.21) --
	(584.30, 44.21) --
	(584.57, 44.21) --
	(584.84, 44.21) --
	(585.11, 44.21) --
	(585.38, 44.21) --
	(585.65, 44.21) --
	(585.92, 44.21) --
	(586.19, 44.21) --
	(586.47, 44.21) --
	(586.74, 44.21) --
	(587.01, 44.21) --
	(587.28, 44.21) --
	(587.55, 44.21) --
	(587.82, 44.21) --
	(588.09, 44.21) --
	(588.36, 44.21) --
	(588.63, 44.21) --
	(588.90, 44.21) --
	(589.18, 44.21) --
	(589.45, 44.21) --
	(589.72, 44.21) --
	(589.99, 44.21) --
	(590.26, 44.21) --
	(590.53, 44.21) --
	(590.80, 44.21) --
	(591.07, 44.21) --
	(591.34, 44.21) --
	(591.61, 44.21) --
	(591.89, 44.21) --
	(592.16, 44.21) --
	(592.43, 44.21) --
	(592.70, 44.21) --
	(592.97, 44.21) --
	(593.24, 44.21) --
	(593.51, 44.21) --
	(593.78, 44.21) --
	(594.05, 44.21) --
	(594.32, 44.21) --
	(594.59, 44.21) --
	(594.87, 44.21) --
	(595.14, 44.21) --
	(595.41, 44.21) --
	(595.68, 44.21) --
	(595.95, 44.21) --
	(596.22, 44.21) --
	(596.49, 44.21) --
	(596.76, 44.21) --
	(597.03, 44.21) --
	(597.30, 44.21) --
	(597.58, 44.21) --
	(597.85, 44.21) --
	(598.12, 44.21) --
	(598.39, 44.21) --
	(598.66, 44.21) --
	(598.93, 44.21) --
	(599.20, 44.21) --
	(599.47, 44.21) --
	(599.74, 44.21) --
	(600.01, 44.21) --
	(600.29, 44.21) --
	(600.56, 44.21) --
	(600.83, 44.21) --
	(601.10, 44.21) --
	(601.37, 44.21) --
	(601.64, 44.21) --
	(601.91, 44.21) --
	(602.18, 44.21) --
	(602.45, 44.21) --
	(602.72, 44.21) --
	(603.00, 44.21) --
	(603.27, 44.21) --
	(603.54, 44.21) --
	(603.81, 44.21) --
	(604.08, 44.21) --
	(604.35, 44.21) --
	(604.62, 44.21) --
	(604.89, 44.21) --
	(605.16, 44.21) --
	(605.43, 44.21) --
	(605.70, 44.21) --
	(605.98, 44.21) --
	(606.25, 44.21) --
	(606.52, 44.21) --
	(606.79, 44.21) --
	(607.06, 44.21) --
	(607.33, 44.21) --
	(607.60, 44.21) --
	(607.87, 44.21) --
	(608.14, 44.21) --
	(608.41, 44.21) --
	(608.69, 44.21) --
	(608.96, 44.21) --
	(609.23, 44.21) --
	(609.50, 44.21) --
	(609.77, 44.21) --
	(610.04, 44.21) --
	(610.31, 44.21) --
	(610.58, 44.21) --
	(610.85, 44.21) --
	(611.12, 44.21) --
	(611.40, 44.21) --
	(611.67, 44.21) --
	(611.94, 44.21) --
	(612.21, 44.21) --
	(612.48, 44.21) --
	(612.75, 44.21) --
	(613.02, 44.21) --
	(613.29, 44.21) --
	(613.56, 44.21) --
	(613.83, 44.21) --
	(614.11, 44.21) --
	(614.38, 44.21) --
	(614.65, 44.21) --
	(614.92, 44.21) --
	(615.19, 44.21) --
	(615.46, 44.21) --
	(615.73, 44.21) --
	(616.00, 44.21) --
	(616.27, 44.21) --
	(616.54, 44.21) --
	(616.81, 44.21) --
	(617.09, 44.21) --
	(617.36, 44.21) --
	(617.63, 44.21) --
	(617.90, 44.21) --
	(618.17, 44.21) --
	(618.44, 44.21) --
	(618.71, 44.21) --
	(618.98, 44.21) --
	(619.25, 44.21) --
	(619.52, 44.21) --
	(619.80, 44.21) --
	(620.07, 44.21) --
	(620.34, 44.21) --
	(620.61, 44.21) --
	(620.88, 44.21) --
	(621.15, 44.21) --
	(621.42, 44.21) --
	(621.69, 44.21) --
	(621.96, 44.21) --
	(622.23, 44.21) --
	(622.51, 44.21) --
	(622.78, 44.21) --
	(623.05, 44.21) --
	(623.32, 44.21) --
	(623.59, 44.21) --
	(623.86, 44.21) --
	(624.13, 44.21) --
	(624.40, 44.21) --
	(624.67, 44.21) --
	(624.94, 44.21) --
	(625.22, 44.21) --
	(625.49, 44.21) --
	(625.76, 44.21) --
	(626.03, 44.21) --
	(626.30, 44.21) --
	(626.57, 44.21) --
	(626.84, 44.21) --
	(627.11, 44.21) --
	(627.38, 44.21) --
	(627.65, 44.21) --
	(627.92, 44.21) --
	(628.20, 44.21) --
	(628.47, 44.21) --
	(628.74, 44.21) --
	(629.01, 44.21) --
	(629.28, 44.21) --
	(629.55, 44.21) --
	(629.82, 44.21) --
	(630.09, 44.21) --
	(630.36, 44.21) --
	(630.63, 44.21) --
	(630.91, 44.21) --
	(631.18, 44.21) --
	(631.45, 44.21) --
	(631.72, 44.21) --
	(631.99, 44.21) --
	(632.26, 44.21) --
	(632.53, 44.21) --
	(632.80, 44.21) --
	(633.07, 44.21) --
	(633.34, 44.21) --
	(633.62, 44.21) --
	(633.89, 44.21) --
	(634.16, 44.21) --
	(634.43, 44.21) --
	(634.70, 44.21) --
	(634.97, 44.21) --
	(635.24, 44.21) --
	(635.51, 44.21) --
	(635.78, 44.21) --
	(636.05, 44.21) --
	(636.33, 44.21) --
	(636.60, 44.21) --
	(636.87, 44.21) --
	(637.14, 44.21) --
	(637.41, 44.21) --
	(637.68, 44.21) --
	(637.95, 44.21) --
	(638.22, 44.21) --
	(638.49, 44.21) --
	(638.76, 44.21) --
	(639.03, 44.21) --
	(639.31, 44.21) --
	(639.58, 44.21) --
	(639.85, 44.21) --
	(640.12, 44.21) --
	(640.39, 44.21) --
	(640.66, 44.21) --
	(640.93, 44.21) --
	(641.20, 44.21) --
	(641.47, 44.21) --
	(641.74, 44.21) --
	(642.02, 44.21) --
	(642.29, 44.21) --
	(642.56, 44.21) --
	(642.83, 44.21) --
	(643.10, 44.21) --
	(643.37, 44.21) --
	(643.64, 44.21) --
	(643.91, 44.21) --
	(644.18, 44.21) --
	(644.45, 44.21) --
	(644.73, 44.21) --
	(645.00, 44.21) --
	(645.27, 44.21) --
	(645.54, 44.21) --
	(645.81, 44.21) --
	(646.08, 44.21) --
	(646.35, 44.21) --
	(646.62, 44.21) --
	(646.89, 44.21) --
	(647.16, 44.21) --
	(647.44, 44.21) --
	(647.71, 44.21) --
	(647.98, 44.21) --
	(648.25, 44.21) --
	(648.52, 44.21) --
	(648.79, 44.21) --
	(649.06, 44.21) --
	(649.33, 44.21) --
	(649.60, 44.21) --
	(649.87, 44.21) --
	(650.15, 44.21) --
	(650.42, 44.21) --
	(650.69, 44.21) --
	(650.96, 44.21) --
	(651.23, 44.21) --
	(651.50, 44.21) --
	(651.77, 44.21) --
	(652.04, 44.21) --
	(652.31, 44.21) --
	(652.58, 44.21) --
	(652.85, 44.21) --
	(653.13, 44.21) --
	(653.40, 44.21) --
	(653.67, 44.21) --
	(653.94, 44.21) --
	(654.21, 44.21) --
	(654.48, 44.21) --
	(654.75, 44.21) --
	(655.02, 44.21) --
	(655.29, 44.21) --
	(655.56, 44.21) --
	(655.84, 44.21) --
	(656.11, 44.21) --
	(656.38, 44.21) --
	(656.65, 44.21) --
	(656.92, 44.21) --
	(657.19, 44.21) --
	(657.46, 44.21) --
	(657.73, 44.21) --
	(658.00, 44.21) --
	(658.27, 44.21) --
	(658.55, 44.21) --
	(658.82, 44.21) --
	(659.09, 44.21) --
	(659.36, 44.21) --
	(659.63, 44.21) --
	(659.90, 44.21) --
	(660.17, 44.21) --
	(660.44, 44.21) --
	(660.71, 44.21) --
	(660.98, 44.21) --
	(661.26, 44.21) --
	(661.53, 44.21) --
	(661.80, 44.21) --
	(662.07, 44.21) --
	(662.34, 44.21) --
	(662.61, 44.21) --
	(662.88, 44.21) --
	(663.15, 44.21) --
	(663.42, 44.21) --
	(663.69, 44.21) --
	(663.96, 44.21) --
	(664.24, 44.21) --
	(664.51, 44.21) --
	(664.78, 44.21) --
	(665.05, 44.21) --
	(665.32, 44.21) --
	(665.59, 44.21) --
	(665.86, 44.21) --
	(666.13, 44.21) --
	(666.40, 44.21) --
	(666.67, 44.21) --
	(666.95, 44.21) --
	(667.22, 44.21) --
	(667.49, 44.21) --
	(667.76, 44.21) --
	(668.03, 44.21) --
	(668.30, 44.21) --
	(668.57, 44.21) --
	(668.84, 44.21) --
	(669.11, 44.21) --
	(669.38, 44.21) --
	(669.66, 44.21) --
	(669.93, 44.21) --
	(670.20, 44.21) --
	(670.47, 44.21) --
	(670.74, 44.21) --
	(671.01, 44.21) --
	(671.28, 44.21) --
	(671.55, 44.21) --
	(671.82, 44.21) --
	(672.09, 44.21) --
	(672.37, 44.21) --
	(672.64, 44.21) --
	(672.91, 44.21) --
	(673.18, 44.21) --
	(673.45, 44.21) --
	(673.72, 44.21) --
	(673.99, 44.21) --
	(674.26, 44.21) --
	(674.53, 44.21) --
	(674.80, 44.21) --
	(675.07, 44.21) --
	(675.35, 44.21) --
	(675.62, 44.21) --
	(675.89, 44.21) --
	(676.16, 44.21) --
	(676.43, 44.21) --
	(676.70, 44.21) --
	(676.97, 44.21) --
	(677.24, 44.21) --
	(677.51, 44.21) --
	(677.78, 44.21) --
	(678.06, 44.21) --
	(678.33, 44.21) --
	(678.60, 44.21) --
	(678.87, 44.21) --
	(679.14, 44.21) --
	(679.41, 44.21) --
	(679.68, 44.21) --
	(679.95, 44.21) --
	(680.22, 44.21) --
	(680.49, 44.21) --
	(680.77, 44.21) --
	(681.04, 44.21) --
	(681.31, 44.21) --
	(681.58, 44.21) --
	(681.85, 44.21) --
	(682.12, 44.21) --
	(682.39, 44.21) --
	(682.66, 44.21) --
	(682.93, 44.21) --
	(683.20, 44.21) --
	(683.48, 44.21) --
	(683.75, 44.21) --
	(684.02, 44.21) --
	(684.29, 44.21) --
	(684.56, 44.21) --
	(684.83, 44.21) --
	(685.10, 44.21) --
	(685.37, 44.21) --
	(685.64, 44.21) --
	(685.91, 44.21) --
	(686.18, 44.21) --
	(686.46, 44.21) --
	(686.73, 44.21) --
	(687.00, 44.21) --
	(687.27, 44.21) --
	(687.54, 44.21) --
	(687.81, 44.21) --
	(688.08, 44.21) --
	(688.35, 44.21) --
	(688.62, 44.21) --
	(688.89, 44.21) --
	(689.17, 44.21) --
	(689.44, 44.21) --
	(689.71, 44.21) --
	(689.98, 44.21) --
	(690.25, 44.21) --
	(690.52, 44.21) --
	(690.79, 44.21) --
	(691.06, 44.21) --
	(691.33, 44.21) --
	(691.60, 44.21) --
	(691.88, 44.21) --
	(692.15, 44.21) --
	(692.42, 44.21) --
	(692.69, 44.21) --
	(692.96, 44.21) --
	(693.23, 44.21) --
	(693.50, 44.21) --
	(693.77, 44.21) --
	(694.04, 44.21) --
	(694.31, 44.21) --
	(694.59, 44.21) --
	(694.86, 44.21) --
	(695.13, 44.21) --
	(695.40, 44.21) --
	(695.67, 44.21) --
	(695.94, 44.21) --
	(696.21, 44.21) --
	(696.48, 44.21) --
	(696.75, 44.21) --
	(697.02, 44.21) --
	(697.29, 44.21) --
	(697.57, 44.21) --
	(697.84, 44.21) --
	(698.11, 44.21) --
	(698.38, 44.21) --
	(698.65, 44.21) --
	(698.92, 44.21) --
	(699.19, 44.21) --
	(699.46, 44.21) --
	(699.73, 44.21) --
	(700.00, 44.21) --
	(700.28, 44.21) --
	(700.55, 44.21) --
	(700.82, 44.21) --
	(701.09, 44.21) --
	(701.36, 44.21) --
	(701.63, 44.21) --
	(701.90, 44.21) --
	(702.17, 44.21) --
	(702.44, 44.21) --
	(702.71, 44.21) --
	(702.99, 44.21) --
	(703.26, 44.21) --
	(703.53, 44.21) --
	(703.80, 44.21) --
	(704.07, 44.21) --
	(704.34, 44.21) --
	(704.61, 44.21) --
	(704.88, 44.21) --
	(705.15, 44.21) --
	(705.42, 44.21) --
	(705.70, 44.21) --
	(705.97, 44.21) --
	(706.24, 44.21) --
	(706.51, 44.21) --
	(706.78, 44.21) --
	(707.05, 44.21) --
	(707.32, 44.21) --
	(707.59, 44.21) --
	(707.86, 44.21) --
	(708.13, 44.21) --
	(708.41, 44.21) --
	(708.68, 44.21) --
	(708.95, 44.21) --
	(709.22, 44.21) --
	(709.49, 44.21) --
	(709.76, 44.21) --
	(709.76, 44.24) --
	(709.49, 44.24) --
	(709.22, 44.24) --
	(708.95, 44.24) --
	(708.68, 44.24) --
	(708.41, 44.24) --
	(708.13, 44.24) --
	(707.86, 44.24) --
	(707.59, 44.24) --
	(707.32, 44.24) --
	(707.05, 44.24) --
	(706.78, 44.24) --
	(706.51, 44.24) --
	(706.24, 44.24) --
	(705.97, 44.24) --
	(705.70, 44.24) --
	(705.42, 44.24) --
	(705.15, 44.24) --
	(704.88, 44.24) --
	(704.61, 44.24) --
	(704.34, 44.24) --
	(704.07, 44.24) --
	(703.80, 44.24) --
	(703.53, 44.23) --
	(703.26, 44.23) --
	(702.99, 44.23) --
	(702.71, 44.23) --
	(702.44, 44.23) --
	(702.17, 44.23) --
	(701.90, 44.23) --
	(701.63, 44.22) --
	(701.36, 44.22) --
	(701.09, 44.22) --
	(700.82, 44.22) --
	(700.55, 44.22) --
	(700.28, 44.22) --
	(700.00, 44.22) --
	(699.73, 44.22) --
	(699.46, 44.22) --
	(699.19, 44.22) --
	(698.92, 44.22) --
	(698.65, 44.22) --
	(698.38, 44.22) --
	(698.11, 44.22) --
	(697.84, 44.22) --
	(697.57, 44.22) --
	(697.29, 44.22) --
	(697.02, 44.23) --
	(696.75, 44.23) --
	(696.48, 44.23) --
	(696.21, 44.23) --
	(695.94, 44.23) --
	(695.67, 44.23) --
	(695.40, 44.23) --
	(695.13, 44.23) --
	(694.86, 44.23) --
	(694.59, 44.24) --
	(694.31, 44.24) --
	(694.04, 44.24) --
	(693.77, 44.24) --
	(693.50, 44.24) --
	(693.23, 44.24) --
	(692.96, 44.24) --
	(692.69, 44.24) --
	(692.42, 44.24) --
	(692.15, 44.24) --
	(691.88, 44.24) --
	(691.60, 44.24) --
	(691.33, 44.24) --
	(691.06, 44.24) --
	(690.79, 44.24) --
	(690.52, 44.24) --
	(690.25, 44.24) --
	(689.98, 44.24) --
	(689.71, 44.25) --
	(689.44, 44.25) --
	(689.17, 44.25) --
	(688.89, 44.25) --
	(688.62, 44.26) --
	(688.35, 44.26) --
	(688.08, 44.27) --
	(687.81, 44.27) --
	(687.54, 44.27) --
	(687.27, 44.28) --
	(687.00, 44.28) --
	(686.73, 44.28) --
	(686.46, 44.29) --
	(686.18, 44.29) --
	(685.91, 44.29) --
	(685.64, 44.29) --
	(685.37, 44.29) --
	(685.10, 44.29) --
	(684.83, 44.29) --
	(684.56, 44.29) --
	(684.29, 44.29) --
	(684.02, 44.29) --
	(683.75, 44.29) --
	(683.48, 44.29) --
	(683.20, 44.29) --
	(682.93, 44.29) --
	(682.66, 44.29) --
	(682.39, 44.29) --
	(682.12, 44.28) --
	(681.85, 44.28) --
	(681.58, 44.28) --
	(681.31, 44.28) --
	(681.04, 44.28) --
	(680.77, 44.28) --
	(680.49, 44.28) --
	(680.22, 44.28) --
	(679.95, 44.28) --
	(679.68, 44.28) --
	(679.41, 44.28) --
	(679.14, 44.28) --
	(678.87, 44.28) --
	(678.60, 44.28) --
	(678.33, 44.28) --
	(678.06, 44.28) --
	(677.78, 44.27) --
	(677.51, 44.27) --
	(677.24, 44.27) --
	(676.97, 44.27) --
	(676.70, 44.27) --
	(676.43, 44.27) --
	(676.16, 44.27) --
	(675.89, 44.27) --
	(675.62, 44.27) --
	(675.35, 44.27) --
	(675.07, 44.27) --
	(674.80, 44.27) --
	(674.53, 44.27) --
	(674.26, 44.27) --
	(673.99, 44.27) --
	(673.72, 44.27) --
	(673.45, 44.27) --
	(673.18, 44.27) --
	(672.91, 44.27) --
	(672.64, 44.27) --
	(672.37, 44.28) --
	(672.09, 44.28) --
	(671.82, 44.28) --
	(671.55, 44.28) --
	(671.28, 44.28) --
	(671.01, 44.28) --
	(670.74, 44.28) --
	(670.47, 44.28) --
	(670.20, 44.28) --
	(669.93, 44.28) --
	(669.66, 44.27) --
	(669.38, 44.27) --
	(669.11, 44.27) --
	(668.84, 44.27) --
	(668.57, 44.27) --
	(668.30, 44.27) --
	(668.03, 44.26) --
	(667.76, 44.26) --
	(667.49, 44.26) --
	(667.22, 44.26) --
	(666.95, 44.26) --
	(666.67, 44.26) --
	(666.40, 44.26) --
	(666.13, 44.26) --
	(665.86, 44.26) --
	(665.59, 44.26) --
	(665.32, 44.26) --
	(665.05, 44.26) --
	(664.78, 44.26) --
	(664.51, 44.26) --
	(664.24, 44.27) --
	(663.96, 44.27) --
	(663.69, 44.27) --
	(663.42, 44.28) --
	(663.15, 44.28) --
	(662.88, 44.28) --
	(662.61, 44.29) --
	(662.34, 44.29) --
	(662.07, 44.29) --
	(661.80, 44.30) --
	(661.53, 44.30) --
	(661.26, 44.31) --
	(660.98, 44.31) --
	(660.71, 44.31) --
	(660.44, 44.32) --
	(660.17, 44.32) --
	(659.90, 44.33) --
	(659.63, 44.33) --
	(659.36, 44.34) --
	(659.09, 44.35) --
	(658.82, 44.35) --
	(658.55, 44.36) --
	(658.27, 44.36) --
	(658.00, 44.37) --
	(657.73, 44.37) --
	(657.46, 44.38) --
	(657.19, 44.38) --
	(656.92, 44.39) --
	(656.65, 44.39) --
	(656.38, 44.40) --
	(656.11, 44.40) --
	(655.84, 44.40) --
	(655.56, 44.40) --
	(655.29, 44.41) --
	(655.02, 44.41) --
	(654.75, 44.41) --
	(654.48, 44.41) --
	(654.21, 44.41) --
	(653.94, 44.41) --
	(653.67, 44.41) --
	(653.40, 44.41) --
	(653.13, 44.41) --
	(652.85, 44.41) --
	(652.58, 44.42) --
	(652.31, 44.42) --
	(652.04, 44.42) --
	(651.77, 44.43) --
	(651.50, 44.43) --
	(651.23, 44.44) --
	(650.96, 44.45) --
	(650.69, 44.46) --
	(650.42, 44.46) --
	(650.15, 44.47) --
	(649.87, 44.49) --
	(649.60, 44.50) --
	(649.33, 44.51) --
	(649.06, 44.52) --
	(648.79, 44.53) --
	(648.52, 44.55) --
	(648.25, 44.56) --
	(647.98, 44.58) --
	(647.71, 44.59) --
	(647.44, 44.61) --
	(647.16, 44.63) --
	(646.89, 44.64) --
	(646.62, 44.66) --
	(646.35, 44.67) --
	(646.08, 44.69) --
	(645.81, 44.71) --
	(645.54, 44.72) --
	(645.27, 44.74) --
	(645.00, 44.76) --
	(644.73, 44.77) --
	(644.45, 44.79) --
	(644.18, 44.80) --
	(643.91, 44.81) --
	(643.64, 44.83) --
	(643.37, 44.84) --
	(643.10, 44.85) --
	(642.83, 44.86) --
	(642.56, 44.87) --
	(642.29, 44.88) --
	(642.02, 44.89) --
	(641.74, 44.90) --
	(641.47, 44.90) --
	(641.20, 44.91) --
	(640.93, 44.92) --
	(640.66, 44.93) --
	(640.39, 44.93) --
	(640.12, 44.94) --
	(639.85, 44.95) --
	(639.58, 44.96) --
	(639.31, 44.97) --
	(639.03, 44.98) --
	(638.76, 44.99) --
	(638.49, 45.00) --
	(638.22, 45.01) --
	(637.95, 45.03) --
	(637.68, 45.04) --
	(637.41, 45.06) --
	(637.14, 45.08) --
	(636.87, 45.10) --
	(636.60, 45.12) --
	(636.33, 45.14) --
	(636.05, 45.16) --
	(635.78, 45.19) --
	(635.51, 45.21) --
	(635.24, 45.24) --
	(634.97, 45.26) --
	(634.70, 45.29) --
	(634.43, 45.31) --
	(634.16, 45.34) --
	(633.89, 45.36) --
	(633.62, 45.39) --
	(633.34, 45.41) --
	(633.07, 45.43) --
	(632.80, 45.45) --
	(632.53, 45.47) --
	(632.26, 45.49) --
	(631.99, 45.51) --
	(631.72, 45.52) --
	(631.45, 45.54) --
	(631.18, 45.55) --
	(630.91, 45.57) --
	(630.63, 45.58) --
	(630.36, 45.60) --
	(630.09, 45.61) --
	(629.82, 45.62) --
	(629.55, 45.64) --
	(629.28, 45.65) --
	(629.01, 45.66) --
	(628.74, 45.68) --
	(628.47, 45.69) --
	(628.20, 45.71) --
	(627.92, 45.73) --
	(627.65, 45.75) --
	(627.38, 45.76) --
	(627.11, 45.78) --
	(626.84, 45.81) --
	(626.57, 45.83) --
	(626.30, 45.85) --
	(626.03, 45.88) --
	(625.76, 45.91) --
	(625.49, 45.94) --
	(625.22, 45.97) --
	(624.94, 46.00) --
	(624.67, 46.04) --
	(624.40, 46.07) --
	(624.13, 46.12) --
	(623.86, 46.16) --
	(623.59, 46.20) --
	(623.32, 46.25) --
	(623.05, 46.30) --
	(622.78, 46.35) --
	(622.51, 46.40) --
	(622.23, 46.45) --
	(621.96, 46.51) --
	(621.69, 46.57) --
	(621.42, 46.63) --
	(621.15, 46.69) --
	(620.88, 46.76) --
	(620.61, 46.82) --
	(620.34, 46.89) --
	(620.07, 46.96) --
	(619.80, 47.03) --
	(619.52, 47.11) --
	(619.25, 47.19) --
	(618.98, 47.27) --
	(618.71, 47.35) --
	(618.44, 47.44) --
	(618.17, 47.53) --
	(617.90, 47.62) --
	(617.63, 47.71) --
	(617.36, 47.80) --
	(617.09, 47.90) --
	(616.81, 47.99) --
	(616.54, 48.08) --
	(616.27, 48.18) --
	(616.00, 48.27) --
	(615.73, 48.35) --
	(615.46, 48.44) --
	(615.19, 48.52) --
	(614.92, 48.60) --
	(614.65, 48.68) --
	(614.38, 48.75) --
	(614.11, 48.82) --
	(613.83, 48.88) --
	(613.56, 48.94) --
	(613.29, 49.00) --
	(613.02, 49.06) --
	(612.75, 49.11) --
	(612.48, 49.16) --
	(612.21, 49.20) --
	(611.94, 49.25) --
	(611.67, 49.29) --
	(611.40, 49.34) --
	(611.12, 49.38) --
	(610.85, 49.43) --
	(610.58, 49.47) --
	(610.31, 49.52) --
	(610.04, 49.57) --
	(609.77, 49.62) --
	(609.50, 49.68) --
	(609.23, 49.74) --
	(608.96, 49.81) --
	(608.69, 49.88) --
	(608.41, 49.95) --
	(608.14, 50.03) --
	(607.87, 50.11) --
	(607.60, 50.19) --
	(607.33, 50.28) --
	(607.06, 50.37) --
	(606.79, 50.46) --
	(606.52, 50.56) --
	(606.25, 50.65) --
	(605.98, 50.75) --
	(605.70, 50.84) --
	(605.43, 50.94) --
	(605.16, 51.03) --
	(604.89, 51.12) --
	(604.62, 51.21) --
	(604.35, 51.29) --
	(604.08, 51.38) --
	(603.81, 51.46) --
	(603.54, 51.53) --
	(603.27, 51.61) --
	(603.00, 51.69) --
	(602.72, 51.76) --
	(602.45, 51.84) --
	(602.18, 51.92) --
	(601.91, 52.00) --
	(601.64, 52.08) --
	(601.37, 52.17) --
	(601.10, 52.26) --
	(600.83, 52.36) --
	(600.56, 52.46) --
	(600.29, 52.57) --
	(600.01, 52.69) --
	(599.74, 52.81) --
	(599.47, 52.93) --
	(599.20, 53.06) --
	(598.93, 53.19) --
	(598.66, 53.32) --
	(598.39, 53.45) --
	(598.12, 53.59) --
	(597.85, 53.72) --
	(597.58, 53.85) --
	(597.30, 53.98) --
	(597.03, 54.10) --
	(596.76, 54.22) --
	(596.49, 54.33) --
	(596.22, 54.44) --
	(595.95, 54.55) --
	(595.68, 54.64) --
	(595.41, 54.73) --
	(595.14, 54.81) --
	(594.87, 54.89) --
	(594.59, 54.95) --
	(594.32, 55.01) --
	(594.05, 55.07) --
	(593.78, 55.11) --
	(593.51, 55.15) --
	(593.24, 55.17) --
	(592.97, 55.19) --
	(592.70, 55.20) --
	(592.43, 55.20) --
	(592.16, 55.20) --
	(591.89, 55.18) --
	(591.61, 55.15) --
	(591.34, 55.11) --
	(591.07, 55.07) --
	(590.80, 55.01) --
	(590.53, 54.94) --
	(590.26, 54.87) --
	(589.99, 54.78) --
	(589.72, 54.69) --
	(589.45, 54.58) --
	(589.18, 54.47) --
	(588.90, 54.36) --
	(588.63, 54.23) --
	(588.36, 54.10) --
	(588.09, 53.96) --
	(587.82, 53.81) --
	(587.55, 53.66) --
	(587.28, 53.50) --
	(587.01, 53.33) --
	(586.74, 53.16) --
	(586.47, 52.97) --
	(586.19, 52.78) --
	(585.92, 52.59) --
	(585.65, 52.38) --
	(585.38, 52.17) --
	(585.11, 51.95) --
	(584.84, 51.72) --
	(584.57, 51.49) --
	(584.30, 51.25) --
	(584.03, 51.01) --
	(583.76, 50.76) --
	(583.48, 50.51) --
	(583.21, 50.25) --
	(582.94, 50.00) --
	(582.67, 49.74) --
	(582.40, 49.49) --
	(582.13, 49.23) --
	(581.86, 48.98) --
	(581.59, 48.74) --
	(581.32, 48.50) --
	(581.05, 48.27) --
	(580.77, 48.04) --
	(580.50, 47.82) --
	(580.23, 47.62) --
	(579.96, 47.41) --
	(579.69, 47.22) --
	(579.42, 47.04) --
	(579.15, 46.86) --
	(578.88, 46.69) --
	(578.61, 46.54) --
	(578.34, 46.39) --
	(578.07, 46.25) --
	(577.79, 46.11) --
	(577.52, 45.98) --
	(577.25, 45.86) --
	(576.98, 45.75) --
	(576.71, 45.64) --
	(576.44, 45.54) --
	(576.17, 45.44) --
	(575.90, 45.35) --
	(575.63, 45.26) --
	(575.36, 45.18) --
	(575.08, 45.10) --
	(574.81, 45.03) --
	(574.54, 44.96) --
	(574.27, 44.89) --
	(574.00, 44.83) --
	(573.73, 44.77) --
	(573.46, 44.72) --
	(573.19, 44.67) --
	(572.92, 44.62) --
	(572.65, 44.58) --
	(572.37, 44.54) --
	(572.10, 44.50) --
	(571.83, 44.47) --
	(571.56, 44.44) --
	(571.29, 44.41) --
	(571.29, 44.21) --
	cycle;
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (553.52, 31.02) rectangle (717.20,211.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57, 41.18) {Pooled estimate};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57, 63.17) {Study 1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57, 85.15) {Study 2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,107.14) {Study 3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,129.13) {Study 4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,151.11) {Study 5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,173.10) {Study 6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,195.09) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77, 44.21) --
	(553.52, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77, 66.20) --
	(553.52, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77, 88.18) --
	(553.52, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,110.17) --
	(553.52,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,132.16) --
	(553.52,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,154.14) --
	(553.52,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,176.13) --
	(553.52,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,198.12) --
	(553.52,198.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (560.96, 28.27) --
	(560.96, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.67, 28.27) --
	(597.67, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.38, 28.27) --
	(634.38, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (671.09, 28.27) --
	(671.09, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (707.80, 28.27) --
	(707.80, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (560.96, 20.01) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (597.67, 20.01) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (634.38, 20.01) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (671.09, 20.01) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (707.80, 20.01) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (635.36,  7.82) {Generation-interval dispersion};
\end{scope}
\end{tikzpicture}

\end{document}