### 2020-05-07 figure2.tex

% Created by tikzDevice version 0.12.3 on 2020-05-07 17:13:04
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 49.68, 44.63) rectangle (281.08,256.12);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.68, 44.63) rectangle (281.08,256.12);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(281.08, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 85.31) --
	(281.08, 85.31);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,125.98) --
	(281.08,125.98);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,166.65) --
	(281.08,166.65);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,207.32) --
	(281.08,207.32);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,247.99) --
	(281.08,247.99);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.38,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (120.88,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (165.38,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (209.88,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.38,123.50) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
	( 76.38,126.52);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
	( 76.38,126.52);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
	( 76.38,126.52);

\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,228.77) --
	(120.88,228.77);

\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,228.77) --
	(120.88, 65.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88, 65.26) --
	(120.88, 65.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,154.56) --
	(165.38,154.56);

\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,154.56) --
	(165.38,104.90);

\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,104.90) --
	(165.38,104.90);

\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,164.33) --
	(209.88,164.33);

\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,164.33) --
	(209.88, 76.82);

\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88, 76.82) --
	(209.88, 76.82);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,252.55) --
	(254.38,252.55);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,252.55) --
	(254.38, 51.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38, 51.71) --
	(254.38, 51.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	( 49.68,256.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 80.90) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,121.57) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,162.24) {3.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,202.91) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,243.58) {4.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 85.31) --
	( 49.68, 85.31);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,125.98) --
	( 49.68,125.98);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,166.65) --
	( 49.68,166.65);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,207.32) --
	( 49.68,207.32);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,247.99) --
	( 49.68,247.99);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(281.08, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 76.38, 40.63) --
	( 76.38, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (120.88, 40.63) --
	(120.88, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (165.38, 40.63) --
	(165.38, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (209.88, 40.63) --
	(209.88, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (254.38, 40.63) --
	(254.38, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 76.38, 28.62) {none};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (120.88, 28.62) {$\mu_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (165.38, 28.62) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (209.88, 28.62) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (254.38, 28.62) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (165.38, 11.11) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,150.38) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at ( 49.68,267.86) {A. Baseline};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] (338.76, 44.63) rectangle (570.16,256.12);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (338.76, 44.63) rectangle (570.16,256.12);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 44.63) --
	(570.16, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 85.31) --
	(570.16, 85.31);

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,125.98) --
	(570.16,125.98);

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,166.65) --
	(570.16,166.65);

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,207.32) --
	(570.16,207.32);

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,247.99) --
	(570.16,247.99);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (365.46,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (409.96,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (454.46,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (498.96,126.52) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (543.46,125.55) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,126.52) --
	(365.46,126.52);

\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,126.52) --
	(365.46,126.52);

\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,126.52) --
	(365.46,126.52);

\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,150.41) --
	(409.96,150.41);

\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,150.41) --
	(409.96,110.38);

\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,110.38) --
	(409.96,110.38);

\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,154.56) --
	(454.46,154.56);

\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,154.56) --
	(454.46,104.90);

\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,104.90) --
	(454.46,104.90);

\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,164.33) --
	(498.96,164.33);

\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,164.33) --
	(498.96, 76.82);

\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96, 76.82) --
	(498.96, 76.82);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,189.36) --
	(543.46,189.36);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,189.36) --
	(543.46, 73.14);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46, 73.14) --
	(543.46, 73.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 44.63) --
	(338.76,256.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56, 40.23) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56, 80.90) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,121.57) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,162.24) {3.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,202.91) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,243.58) {4.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76, 44.63) --
	(338.76, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76, 85.31) --
	(338.76, 85.31);

\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,125.98) --
	(338.76,125.98);

\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,166.65) --
	(338.76,166.65);

\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,207.32) --
	(338.76,207.32);

\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,247.99) --
	(338.76,247.99);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 44.63) --
	(570.16, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (365.46, 40.63) --
	(365.46, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (409.96, 40.63) --
	(409.96, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (454.46, 40.63) --
	(454.46, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (498.96, 40.63) --
	(498.96, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (543.46, 40.63) --
	(543.46, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (365.46, 28.62) {none};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (409.96, 28.62) {$\hat{\mu}_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (454.46, 28.62) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (498.96, 28.62) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (543.46, 28.62) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (454.46, 11.11) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (308.10,150.38) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at (338.76,267.86) {B. Reduced uncertainty in $r$};
\end{scope}
\end{tikzpicture}

\end{document}