### 2020-05-07 compare_R0.tex

% Created by tikzDevice version 0.12.3 on 2020-05-07 17:27:07
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 18.22) rectangle (572.66,247.13);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 18.22) rectangle (572.66,247.13);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 25.76) --
	(572.66, 25.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 79.98) --
	(572.66, 79.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,134.20) --
	(572.66,134.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,188.41) --
	(572.66,188.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,242.63) --
	(572.66,242.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 52.87) --
	(572.66, 52.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,107.09) --
	(572.66,107.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,161.31) --
	(572.66,161.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,215.52) --
	(572.66,215.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 18.22) --
	( 67.22,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.72, 18.22) --
	(133.72,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.23, 18.22) --
	(200.23,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (266.73, 18.22) --
	(266.73,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.24, 18.22) --
	(333.24,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.75, 18.22) --
	(399.75,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.25, 18.22) --
	(466.25,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.76, 18.22) --
	(532.76,247.13);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 55.23) -- (572.66, 55.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,122.16) -- (572.66,122.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 75.59) rectangle ( 70.79, 82.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 75.59) -- ( 70.79, 82.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 82.72) -- ( 70.79, 75.59);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (141.15, 78.87) -- (151.24, 78.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.19, 73.83) -- (146.19, 83.92);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (134.31, 56.77) --
	(141.45, 56.77) --
	(141.45, 63.91) --
	(134.31, 63.91) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (129.57, 78.05) --
	(134.37, 69.73) --
	(124.76, 69.73) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (121.25, 66.37) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.65, 66.49) -- (217.74, 66.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.70, 61.44) -- (212.70, 71.54);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (200.82, 65.00) --
	(207.95, 65.00) --
	(207.95, 72.13) --
	(200.82, 72.13) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (196.07, 84.79) --
	(200.88, 76.46) --
	(191.27, 76.46) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (187.76, 67.91) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.16, 66.66) -- (284.25, 66.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.20, 61.62) -- (279.20, 71.71);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (267.32, 74.74) --
	(274.46, 74.74) --
	(274.46, 81.88) --
	(267.32, 81.88) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (262.58,100.58) --
	(267.38, 92.26) --
	(257.77, 92.26) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (254.26, 77.41) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.66,101.40) -- (350.76,101.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.71, 96.36) -- (345.71,106.45);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (333.83,110.42) --
	(340.96,110.42) --
	(340.96,117.56) --
	(333.83,117.56) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (329.08, 78.47) --
	(333.89, 70.15) --
	(324.28, 70.15) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (320.77,101.64) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (407.17, 55.76) -- (417.26, 55.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.22, 50.72) -- (412.22, 60.81);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (400.33, 46.35) --
	(407.47, 46.35) --
	(407.47, 53.48) --
	(400.33, 53.48) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (395.59,101.86) --
	(400.39, 93.54) --
	(390.78, 93.54) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (387.28, 56.59) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.67,105.92) -- (483.77,105.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.72,100.87) -- (478.72,110.97);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (466.84,156.18) --
	(473.98,156.18) --
	(473.98,163.32) --
	(466.84,163.32) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (462.09, 95.96) --
	(466.90, 87.64) --
	(457.29, 87.64) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (453.78,145.76) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (540.18, 55.35) -- (550.27, 55.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.23, 50.30) -- (545.23, 60.40);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (533.34, 66.58) --
	(540.48, 66.58) --
	(540.48, 73.72) --
	(533.34, 73.72) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (528.60,111.22) --
	(533.41,102.89) --
	(523.79,102.89) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (520.29, 66.12) circle (  3.57);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,122.16) --
	( 67.22,122.16);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,122.16) --
	( 67.22, 55.23);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22, 55.23) --
	( 67.22, 55.23);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19,148.57) --
	(146.19,148.57);

\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19,148.57) --
	(146.19, 42.52);

\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19, 42.52) --
	(146.19, 42.52);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 83.60) --
	(137.88, 83.60);

\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 83.60) --
	(137.88, 38.44);

\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 38.44) --
	(137.88, 38.44);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 93.55) --
	(129.57, 93.55);

\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 93.55) --
	(129.57, 57.68);

\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 57.68) --
	(129.57, 57.68);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 92.13) --
	(121.25, 92.13);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 92.13) --
	(121.25, 40.71);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 40.71) --
	(121.25, 40.71);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70,109.43) --
	(212.70,109.43);

\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70,109.43) --
	(212.70, 40.25);

\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70, 40.25) --
	(212.70, 40.25);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38,108.47) --
	(204.38,108.47);

\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38,108.47) --
	(204.38, 40.60);

\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38, 40.60) --
	(204.38, 40.60);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07,112.61) --
	(196.07,112.61);

\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07,112.61) --
	(196.07, 59.21);

\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07, 59.21) --
	(196.07, 59.21);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76,105.16) --
	(187.76,105.16);

\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76,105.16) --
	(187.76, 40.54);

\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76, 40.54) --
	(187.76, 40.54);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 84.95) --
	(279.20, 84.95);

\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 84.95) --
	(279.20, 52.73);

\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 52.73) --
	(279.20, 52.73);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89,103.19) --
	(270.89,103.19);

\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89,103.19) --
	(270.89, 59.79);

\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89, 59.79) --
	(270.89, 59.79);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58,154.24) --
	(262.58,154.24);

\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58,154.24) --
	(262.58, 65.38);

\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58, 65.38) --
	(262.58, 65.38);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 97.96) --
	(254.26, 97.96);

\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 97.96) --
	(254.26, 60.10);

\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 60.10) --
	(254.26, 60.10);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71,126.97) --
	(345.71,126.97);

\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71,126.97) --
	(345.71, 74.24);

\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71, 74.24) --
	(345.71, 74.24);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,131.59) --
	(337.40,131.59);

\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,131.59) --
	(337.40,100.29);

\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,100.29) --
	(337.40,100.29);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08,101.76) --
	(329.08,101.76);

\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08,101.76) --
	(329.08, 55.45);

\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08, 55.45) --
	(329.08, 55.45);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77,107.13) --
	(320.77,107.13);

\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77,107.13) --
	(320.77, 96.34);

\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77, 96.34) --
	(320.77, 96.34);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 98.29) --
	(412.22, 98.29);

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 98.29) --
	(412.22, 29.35);

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 29.35) --
	(412.22, 29.35);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 91.28) --
	(403.90, 91.28);

\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 91.28) --
	(403.90, 28.63);

\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 28.63) --
	(403.90, 28.63);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59,165.53) --
	(395.59,165.53);

\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59,165.53) --
	(395.59, 61.03);

\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59, 61.03) --
	(395.59, 61.03);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 96.20) --
	(387.28, 96.20);

\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 96.20) --
	(387.28, 29.36);

\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 29.36) --
	(387.28, 29.36);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72,148.26) --
	(478.72,148.26);

\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72,148.26) --
	(478.72, 74.69);

\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72, 74.69) --
	(478.72, 74.69);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,223.85) --
	(470.41,223.85);

\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,223.85) --
	(470.41,114.90);

\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,114.90) --
	(470.41,114.90);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09,146.48) --
	(462.09,146.48);

\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09,146.48) --
	(462.09, 62.15);

\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09, 62.15) --
	(462.09, 62.15);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,191.49) --
	(453.78,191.49);

\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,191.49) --
	(453.78,109.22);

\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,109.22) --
	(453.78,109.22);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 67.69) --
	(545.23, 67.69);

\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 67.69) --
	(545.23, 46.39);

\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 46.39) --
	(545.23, 46.39);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 79.88) --
	(536.91, 79.88);

\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 79.88) --
	(536.91, 63.12);

\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 63.12) --
	(536.91, 63.12);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60,236.72) --
	(528.60,236.72);

\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60,236.72) --
	(528.60, 61.88);

\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60, 61.88) --
	(528.60, 61.88);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 82.44) --
	(520.29, 82.44);

\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 82.44) --
	(520.29, 52.97);

\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 52.97) --
	(520.29, 52.97);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 18.22) rectangle (572.66,247.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 49.84) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,104.06) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,158.27) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,212.49) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 52.87) --
	( 27.31, 52.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,107.09) --
	( 27.31,107.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,161.31) --
	( 27.31,161.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,215.52) --
	( 27.31,215.52);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 15.47) --
	( 67.22, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.72, 15.47) --
	(133.72, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.23, 15.47) --
	(200.23, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (266.73, 15.47) --
	(266.73, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.24, 15.47) --
	(333.24, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.75, 15.47) --
	(399.75, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.25, 15.47) --
	(466.25, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.76, 15.47) --
	(532.76, 18.22);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 67.22,  7.21) {Pooled estimate};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (133.72,  7.21) {Study 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (200.23,  7.21) {Study 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (266.73,  7.21) {Study 3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (333.24,  7.21) {Study 4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (399.75,  7.21) {Study 5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (466.25,  7.21) {Study 6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (532.76,  7.21) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,132.67) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 67.57,258.13) rectangle (532.41,283.58);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 78.57,263.63) rectangle ( 93.02,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] ( 85.79,270.85) circle (  3.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 80.01,270.85) -- ( 91.57,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (120.69,263.63) rectangle (135.14,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (127.91,276.40) --
	(132.72,268.08) --
	(123.11,268.08) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (122.13,270.85) -- (133.69,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (238.08,263.63) rectangle (252.53,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (241.74,267.28) --
	(248.87,267.28) --
	(248.87,274.42) --
	(241.74,274.42) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 1.1pt,line join=round] (239.52,270.85) -- (251.09,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (358.89,263.63) rectangle (373.35,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.07,270.85) -- (371.17,270.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.12,265.81) -- (366.12,275.90);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 1.1pt,line join=round] (360.34,270.85) -- (371.90,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (497.67,263.63) rectangle (512.12,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,267.28) rectangle (508.46,274.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,267.28) -- (508.46,274.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,274.42) -- (508.46,267.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (499.11,270.85) -- (510.68,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 98.52,267.82) {base};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.64,267.82) {exponential growth rate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (258.03,267.82) {mean generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (378.85,267.82) {generation-interval dispersion};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (517.62,267.82) {all};
\end{scope}
\end{tikzpicture}

\end{document}