### 2020-05-07 propagate_pix.tex

% Created by tikzDevice version 0.12.3 on 2020-05-07 18:04:06
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (505.89,505.89);
\end{scope}
\begin{scope}
\path[clip] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(497.89, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,127.00) --
	(497.89,127.00);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,209.36) --
	(497.89,209.36);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,291.73) --
	(497.89,291.73);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,374.10) --
	(497.89,374.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,456.46) --
	(497.89,456.46);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.59,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,204.36) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,417.54) --
	(187.59,417.54);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,417.54) --
	(187.59, 86.40);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59, 86.40) --
	(187.59, 86.40);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
	(273.79,267.25);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
	(273.79,166.69);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,166.69) --
	(273.79,166.69);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
	(359.98,287.03);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
	(359.98,109.81);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,109.81) --
	(359.98,109.81);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,465.70) --
	(446.17,465.70);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,465.70) --
	(446.17, 58.97);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17, 58.97) --
	(446.17, 58.97);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	( 49.68,472.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,122.59) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,204.96) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,287.32) {3.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,369.69) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,452.05) {4.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,127.00) --
	( 49.68,127.00);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,209.36) --
	( 49.68,209.36);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,291.73) --
	( 49.68,291.73);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,374.10) --
	( 49.68,374.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,456.46) --
	( 49.68,456.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(497.89, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (101.40, 40.63) --
	(101.40, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (187.59, 40.63) --
	(187.59, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (273.79, 40.63) --
	(273.79, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (359.98, 40.63) --
	(359.98, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (446.17, 40.63) --
	(446.17, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (101.40, 28.62) {none};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (187.59, 28.62) {$\mu_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (273.79, 28.62) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (359.98, 28.62) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (446.17, 28.62) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (273.79, 11.11) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,258.78) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at ( 49.68,484.67) {A. Baseline};
\end{scope}
\end{tikzpicture}
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (505.89,505.89);
\end{scope}
\begin{scope}
\path[clip] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(497.89, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,127.00) --
	(497.89,127.00);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,209.36) --
	(497.89,209.36);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,291.73) --
	(497.89,291.73);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,374.10) --
	(497.89,374.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,456.46) --
	(497.89,456.46);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.59,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,208.51) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,258.85) --
	(187.59,258.85);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,258.85) --
	(187.59,177.77);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,177.77) --
	(187.59,177.77);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
	(273.79,267.25);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
	(273.79,166.69);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,166.69) --
	(273.79,166.69);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
	(359.98,287.03);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
	(359.98,109.81);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,109.81) --
	(359.98,109.81);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,337.72) --
	(446.17,337.72);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,337.72) --
	(446.17,102.36);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,102.36) --
	(446.17,102.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	( 49.68,472.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,122.59) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,204.96) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,287.32) {3.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,369.69) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,452.05) {4.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,127.00) --
	( 49.68,127.00);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,209.36) --
	( 49.68,209.36);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,291.73) --
	( 49.68,291.73);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,374.10) --
	( 49.68,374.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,456.46) --
	( 49.68,456.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(497.89, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (101.40, 40.63) --
	(101.40, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (187.59, 40.63) --
	(187.59, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (273.79, 40.63) --
	(273.79, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (359.98, 40.63) --
	(359.98, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (446.17, 40.63) --
	(446.17, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (101.40, 28.62) {none};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (187.59, 28.62) {$\hat{\mu}_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (273.79, 28.62) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (359.98, 28.62) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (446.17, 28.62) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (273.79, 11.11) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,258.78) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at ( 49.68,484.67) {B. Reduced uncertainty in $r$};
\end{scope}
\end{tikzpicture}

\end{document}