### 2020-06-21 compare_R0.tex
done

% Created by tikzDevice version 0.12.3 on 2020-06-20 18:53:21
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 18.22) rectangle (572.66,232.67);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 18.22) rectangle (572.66,232.67);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 25.29) --
	(572.66, 25.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 76.08) --
	(572.66, 76.08);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,126.87) --
	(572.66,126.87);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,177.67) --
	(572.66,177.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,228.46) --
	(572.66,228.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 50.68) --
	(572.66, 50.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,101.48) --
	(572.66,101.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,152.27) --
	(572.66,152.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,203.06) --
	(572.66,203.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 18.22) --
	( 67.22,232.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.72, 18.22) --
	(133.72,232.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.23, 18.22) --
	(200.23,232.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (266.73, 18.22) --
	(266.73,232.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.24, 18.22) --
	(333.24,232.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.75, 18.22) --
	(399.75,232.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.25, 18.22) --
	(466.25,232.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.76, 18.22) --
	(532.76,232.67);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 52.89) -- (572.66, 52.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,115.60) -- (572.66,115.60);
\definecolor{drawColor}{RGB}{182,117,224}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 71.74) -- ( 70.79, 78.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 78.88) -- ( 70.79, 71.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.17, 75.31) -- ( 72.26, 75.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 67.22, 70.26) -- ( 67.22, 80.35);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.45, 71.47) rectangle (150.59, 78.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.45, 71.47) -- (150.59, 78.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.45, 78.61) -- (150.59, 71.47);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.33, 57.68) -- (145.42, 57.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.37, 52.63) -- (140.37, 62.73);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (130.15, 65.51) --
	(137.29, 65.51) --
	(137.29, 72.64) --
	(130.15, 72.64) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (127.07, 68.88) --
	(131.88, 60.55) --
	(122.27, 60.55) --
	cycle;
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.96, 59.87) rectangle (217.10, 67.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.96, 59.87) -- (217.10, 67.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.96, 67.01) -- (217.10, 59.87);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.83, 65.39) -- (211.93, 65.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.88, 60.34) -- (206.88, 70.43);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (196.66, 71.82) --
	(203.80, 71.82) --
	(203.80, 78.95) --
	(196.66, 78.95) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (193.58, 70.32) --
	(198.38, 62.00) --
	(188.77, 62.00) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (186.93, 65.92) circle (  3.57);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.47, 60.04) rectangle (283.60, 67.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.47, 60.04) -- (283.60, 67.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.47, 67.17) -- (283.60, 60.04);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.34, 74.52) -- (278.43, 74.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.38, 69.47) -- (273.38, 79.56);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (263.17, 86.61) --
	(270.30, 86.61) --
	(270.30, 93.75) --
	(263.17, 93.75) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (260.08, 79.22) --
	(264.89, 70.90) --
	(255.28, 70.90) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (253.43, 74.05) circle (  3.57);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.97, 92.58) rectangle (350.11, 99.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.97, 92.58) -- (350.11, 99.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.97, 99.72) -- (350.11, 92.58);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.84,107.95) -- (344.94,107.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.89,102.90) -- (339.89,112.99);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (329.67, 65.90) --
	(336.81, 65.90) --
	(336.81, 73.04) --
	(329.67, 73.04) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (326.59,101.92) --
	(331.40, 93.60) --
	(321.78, 93.60) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (319.94, 96.40) circle (  3.57);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.48, 49.82) rectangle (416.61, 56.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.48, 49.82) -- (416.61, 56.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.48, 56.96) -- (416.61, 49.82);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.35, 47.91) -- (411.44, 47.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.40, 42.87) -- (406.40, 52.96);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (396.18, 87.82) --
	(403.31, 87.82) --
	(403.31, 94.95) --
	(396.18, 94.95) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (393.09, 59.72) --
	(397.90, 51.39) --
	(388.29, 51.39) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (386.44, 55.76) circle (  3.57);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.98, 96.81) rectangle (483.12,103.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.98, 96.81) -- (483.12,103.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.98,103.95) -- (483.12, 96.81);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.85,150.81) -- (477.95,150.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.90,145.76) -- (472.90,155.86);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (462.68, 82.29) --
	(469.82, 82.29) --
	(469.82, 89.42) --
	(462.68, 89.42) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (459.60,143.26) --
	(464.41,134.93) --
	(454.79,134.93) --
	cycle;
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (452.95,138.81) circle (  3.57);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.49, 49.44) rectangle (549.63, 56.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.49, 49.44) -- (549.63, 56.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.49, 56.57) -- (549.63, 49.44);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.36, 66.87) -- (544.45, 66.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.41, 61.82) -- (539.41, 71.92);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (529.19, 96.58) --
	(536.33, 96.58) --
	(536.33,103.71) --
	(529.19,103.71) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (526.11, 68.64) --
	(530.91, 60.32) --
	(521.30, 60.32) --
	cycle;
\definecolor{drawColor}{RGB}{182,117,224}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,115.60) --
	( 67.22,115.60);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,115.60) --
	( 67.22, 52.89);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22, 52.89) --
	( 67.22, 52.89);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (147.02,140.34) --
	(147.02,140.34);

\path[draw=drawColor,line width= 1.1pt,line join=round] (147.02,140.34) --
	(147.02, 40.98);

\path[draw=drawColor,line width= 1.1pt,line join=round] (147.02, 40.98) --
	(147.02, 40.98);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (140.37, 79.47) --
	(140.37, 79.47);

\path[draw=drawColor,line width= 1.1pt,line join=round] (140.37, 79.47) --
	(140.37, 37.16);

\path[draw=drawColor,line width= 1.1pt,line join=round] (140.37, 37.16) --
	(140.37, 37.16);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (133.72, 88.80) --
	(133.72, 88.80);

\path[draw=drawColor,line width= 1.1pt,line join=round] (133.72, 88.80) --
	(133.72, 55.19);

\path[draw=drawColor,line width= 1.1pt,line join=round] (133.72, 55.19) --
	(133.72, 55.19);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.07, 87.46) --
	(127.07, 87.46);

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.07, 87.46) --
	(127.07, 39.29);

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.07, 39.29) --
	(127.07, 39.29);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (120.42, 88.78) --
	(120.42, 88.78);

\path[draw=drawColor,line width= 1.1pt,line join=round] (120.42, 88.78) --
	(120.42, 37.98);

\path[draw=drawColor,line width= 1.1pt,line join=round] (120.42, 37.98) --
	(120.42, 37.98);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (213.53,103.67) --
	(213.53,103.67);

\path[draw=drawColor,line width= 1.1pt,line join=round] (213.53,103.67) --
	(213.53, 38.86);

\path[draw=drawColor,line width= 1.1pt,line join=round] (213.53, 38.86) --
	(213.53, 38.86);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (206.88,102.77) --
	(206.88,102.77);

\path[draw=drawColor,line width= 1.1pt,line join=round] (206.88,102.77) --
	(206.88, 39.19);

\path[draw=drawColor,line width= 1.1pt,line join=round] (206.88, 39.19) --
	(206.88, 39.19);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (200.23,106.65) --
	(200.23,106.65);

\path[draw=drawColor,line width= 1.1pt,line join=round] (200.23,106.65) --
	(200.23, 56.62);

\path[draw=drawColor,line width= 1.1pt,line join=round] (200.23, 56.62) --
	(200.23, 56.62);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (193.58, 99.67) --
	(193.58, 99.67);

\path[draw=drawColor,line width= 1.1pt,line join=round] (193.58, 99.67) --
	(193.58, 39.13);

\path[draw=drawColor,line width= 1.1pt,line join=round] (193.58, 39.13) --
	(193.58, 39.13);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (186.93, 88.78) --
	(186.93, 88.78);

\path[draw=drawColor,line width= 1.1pt,line join=round] (186.93, 88.78) --
	(186.93, 37.98);

\path[draw=drawColor,line width= 1.1pt,line join=round] (186.93, 37.98) --
	(186.93, 37.98);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (280.03, 80.74) --
	(280.03, 80.74);

\path[draw=drawColor,line width= 1.1pt,line join=round] (280.03, 80.74) --
	(280.03, 50.55);

\path[draw=drawColor,line width= 1.1pt,line join=round] (280.03, 50.55) --
	(280.03, 50.55);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (273.38, 97.82) --
	(273.38, 97.82);

\path[draw=drawColor,line width= 1.1pt,line join=round] (273.38, 97.82) --
	(273.38, 57.16);

\path[draw=drawColor,line width= 1.1pt,line join=round] (273.38, 57.16) --
	(273.38, 57.16);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (266.73,145.65) --
	(266.73,145.65);

\path[draw=drawColor,line width= 1.1pt,line join=round] (266.73,145.65) --
	(266.73, 62.40);

\path[draw=drawColor,line width= 1.1pt,line join=round] (266.73, 62.40) --
	(266.73, 62.40);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (260.08, 92.92) --
	(260.08, 92.92);

\path[draw=drawColor,line width= 1.1pt,line join=round] (260.08, 92.92) --
	(260.08, 57.46);

\path[draw=drawColor,line width= 1.1pt,line join=round] (260.08, 57.46) --
	(260.08, 57.46);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (253.43, 93.10) --
	(253.43, 93.10);

\path[draw=drawColor,line width= 1.1pt,line join=round] (253.43, 93.10) --
	(253.43, 57.79);

\path[draw=drawColor,line width= 1.1pt,line join=round] (253.43, 57.79) --
	(253.43, 57.79);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (346.54,120.10) --
	(346.54,120.10);

\path[draw=drawColor,line width= 1.1pt,line join=round] (346.54,120.10) --
	(346.54, 70.70);

\path[draw=drawColor,line width= 1.1pt,line join=round] (346.54, 70.70) --
	(346.54, 70.70);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (339.89,124.43) --
	(339.89,124.43);

\path[draw=drawColor,line width= 1.1pt,line join=round] (339.89,124.43) --
	(339.89, 95.11);

\path[draw=drawColor,line width= 1.1pt,line join=round] (339.89, 95.11) --
	(339.89, 95.11);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (333.24, 96.49) --
	(333.24, 96.49);

\path[draw=drawColor,line width= 1.1pt,line join=round] (333.24, 96.49) --
	(333.24, 53.10);

\path[draw=drawColor,line width= 1.1pt,line join=round] (333.24, 53.10) --
	(333.24, 53.10);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (326.59,101.51) --
	(326.59,101.51);

\path[draw=drawColor,line width= 1.1pt,line join=round] (326.59,101.51) --
	(326.59, 91.41);

\path[draw=drawColor,line width= 1.1pt,line join=round] (326.59, 91.41) --
	(326.59, 91.41);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (319.94,101.48) --
	(319.94,101.48);

\path[draw=drawColor,line width= 1.1pt,line join=round] (319.94,101.48) --
	(319.94, 91.32);

\path[draw=drawColor,line width= 1.1pt,line join=round] (319.94, 91.32) --
	(319.94, 91.32);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (413.05, 93.24) --
	(413.05, 93.24);

\path[draw=drawColor,line width= 1.1pt,line join=round] (413.05, 93.24) --
	(413.05, 28.65);

\path[draw=drawColor,line width= 1.1pt,line join=round] (413.05, 28.65) --
	(413.05, 28.65);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (406.40, 86.66) --
	(406.40, 86.66);

\path[draw=drawColor,line width= 1.1pt,line join=round] (406.40, 86.66) --
	(406.40, 27.97);

\path[draw=drawColor,line width= 1.1pt,line join=round] (406.40, 27.97) --
	(406.40, 27.97);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (399.75,156.23) --
	(399.75,156.23);

\path[draw=drawColor,line width= 1.1pt,line join=round] (399.75,156.23) --
	(399.75, 58.32);

\path[draw=drawColor,line width= 1.1pt,line join=round] (399.75, 58.32) --
	(399.75, 58.32);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (393.09, 91.27) --
	(393.09, 91.27);

\path[draw=drawColor,line width= 1.1pt,line join=round] (393.09, 91.27) --
	(393.09, 28.65);

\path[draw=drawColor,line width= 1.1pt,line join=round] (393.09, 28.65) --
	(393.09, 28.65);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (386.44, 96.40) --
	(386.44, 96.40);

\path[draw=drawColor,line width= 1.1pt,line join=round] (386.44, 96.40) --
	(386.44, 35.44);

\path[draw=drawColor,line width= 1.1pt,line join=round] (386.44, 35.44) --
	(386.44, 35.44);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (479.55,140.05) --
	(479.55,140.05);

\path[draw=drawColor,line width= 1.1pt,line join=round] (479.55,140.05) --
	(479.55, 71.12);

\path[draw=drawColor,line width= 1.1pt,line join=round] (479.55, 71.12) --
	(479.55, 71.12);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (472.90,210.87) --
	(472.90,210.87);

\path[draw=drawColor,line width= 1.1pt,line join=round] (472.90,210.87) --
	(472.90,108.80);

\path[draw=drawColor,line width= 1.1pt,line join=round] (472.90,108.80) --
	(472.90,108.80);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (466.25,138.38) --
	(466.25,138.38);

\path[draw=drawColor,line width= 1.1pt,line join=round] (466.25,138.38) --
	(466.25, 59.37);

\path[draw=drawColor,line width= 1.1pt,line join=round] (466.25, 59.37) --
	(466.25, 59.37);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (459.60,180.55) --
	(459.60,180.55);

\path[draw=drawColor,line width= 1.1pt,line join=round] (459.60,180.55) --
	(459.60,103.48);

\path[draw=drawColor,line width= 1.1pt,line join=round] (459.60,103.48) --
	(459.60,103.48);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (452.95,180.21) --
	(452.95,180.21);

\path[draw=drawColor,line width= 1.1pt,line join=round] (452.95,180.21) --
	(452.95,105.54);

\path[draw=drawColor,line width= 1.1pt,line join=round] (452.95,105.54) --
	(452.95,105.54);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (546.06, 64.57) --
	(546.06, 64.57);

\path[draw=drawColor,line width= 1.1pt,line join=round] (546.06, 64.57) --
	(546.06, 44.61);

\path[draw=drawColor,line width= 1.1pt,line join=round] (546.06, 44.61) --
	(546.06, 44.61);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (539.41, 75.99) --
	(539.41, 75.99);

\path[draw=drawColor,line width= 1.1pt,line join=round] (539.41, 75.99) --
	(539.41, 60.28);

\path[draw=drawColor,line width= 1.1pt,line join=round] (539.41, 60.28) --
	(539.41, 60.28);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (532.76,222.92) --
	(532.76,222.92);

\path[draw=drawColor,line width= 1.1pt,line join=round] (532.76,222.92) --
	(532.76, 59.12);

\path[draw=drawColor,line width= 1.1pt,line join=round] (532.76, 59.12) --
	(532.76, 59.12);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (526.11, 78.38) --
	(526.11, 78.38);

\path[draw=drawColor,line width= 1.1pt,line join=round] (526.11, 78.38) --
	(526.11, 50.78);

\path[draw=drawColor,line width= 1.1pt,line join=round] (526.11, 50.78) --
	(526.11, 50.78);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (519.46, 78.62) --
	(519.46, 78.62);

\path[draw=drawColor,line width= 1.1pt,line join=round] (519.46, 78.62) --
	(519.46, 50.68);

\path[draw=drawColor,line width= 1.1pt,line join=round] (519.46, 50.68) --
	(519.46, 50.68);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 18.22) rectangle (572.66,232.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 47.65) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 98.45) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,149.24) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,200.03) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 50.68) --
	( 27.31, 50.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,101.48) --
	( 27.31,101.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,152.27) --
	( 27.31,152.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,203.06) --
	( 27.31,203.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 15.47) --
	( 67.22, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.72, 15.47) --
	(133.72, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.23, 15.47) --
	(200.23, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (266.73, 15.47) --
	(266.73, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.24, 15.47) --
	(333.24, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.75, 15.47) --
	(399.75, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.25, 15.47) --
	(466.25, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.76, 15.47) --
	(532.76, 18.22);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 67.22,  7.21) {Pooled estimate};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (133.72,  7.21) {Study 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (200.23,  7.21) {Study 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (266.73,  7.21) {Study 3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (333.24,  7.21) {Study 4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (399.75,  7.21) {Study 5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (466.25,  7.21) {Study 6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (532.76,  7.21) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,125.45) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (135.69,243.67) rectangle (464.28,283.58);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (146.69,263.63) rectangle (161.14,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (153.92,270.85) circle (  3.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (148.13,270.85) -- (159.70,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (146.69,249.17) rectangle (161.14,263.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (153.92,261.95) --
	(158.72,253.62) --
	(149.11,253.62) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 1.1pt,line join=round] (148.13,256.40) -- (159.70,256.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (204.70,263.63) rectangle (219.15,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{170,144,0}

\path[fill=fillColor] (208.35,267.28) --
	(215.49,267.28) --
	(215.49,274.42) --
	(208.35,274.42) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{170,144,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (206.14,270.85) -- (217.70,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (204.70,249.17) rectangle (219.15,263.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.88,256.40) -- (216.97,256.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.92,251.35) -- (211.92,261.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,170,90}

\path[draw=drawColor,line width= 1.1pt,line join=round] (206.14,256.40) -- (217.70,256.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (325.51,263.63) rectangle (339.96,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,267.28) rectangle (336.30,274.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,267.28) -- (336.30,274.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,274.42) -- (336.30,267.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,166,202}

\path[draw=drawColor,line width= 1.1pt,line join=round] (326.95,270.85) -- (338.52,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (325.51,249.17) rectangle (339.96,263.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{182,117,224}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,252.83) -- (336.30,259.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,259.97) -- (336.30,252.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (327.69,256.40) -- (337.78,256.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (332.74,251.35) -- (332.74,261.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{182,117,224}

\path[draw=drawColor,line width= 1.1pt,line join=round] (326.95,256.40) -- (338.52,256.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.64,267.82) {reported};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.64,253.37) {base};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (224.65,267.82) {exponential growth rate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (224.65,253.37) {mean generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.46,267.82) {generation-interval dispersion};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.46,253.37) {all};
\end{scope}
\end{tikzpicture}

\end{document}